### 2023-10-13 ms_preprint/ms_preprint.tex

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Seabird assemblages are linked to the major western boundary
current off eastern Australia}

\author{
    Nicholas W. Daudt
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{\href{mailto:nicholaswdaudt@gmail.com}{\nolinkurl{nicholaswdaudt@gmail.com}}} \\
   \And
    Eric J. Woehler
   \\
    Australasian Seabird Group \\
    BirdLife Australia \\
  Hobart, Australia \\
  \texttt{} \\
   \And
    Matthew R. Schofield
   \\
    Department of Mathematics and Statistics \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
   \And
    Robert O. Smith
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
   \And
    Leandro Bugoni
   \\
    Institute of Biological Sciences \\
    Universidade Federal do Rio Grande--FURG \\
  Rio Grande, RS, Brazil \\
  \texttt{} \\
   \And
    William J. Rayment
   \\
    Department of Marine Science \\
    University of Otago \\
  Dunedin, Aotearoa New Zealand \\
  \texttt{} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{setspace}\doublespacing
\usepackage{pdflscape}
\DeclareUnicodeCharacter{0394}{\ensuremath{\Delta}}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
\textbf{Aim}: To identify assemblages of seabird species and relate them
to environmental conditions off eastern Australia, a world seabird
hotspot. \textbf{Methods}: We used 15 seabird at-sea ship-based survey
data collected over 37° of latitude, from 2016 to 2021, and fitted
Region of Common Profile (RCP) mixture models for each season and type
of data (presence-absence and abundance). RCP groups are defined as a
set of regions where the probability of encountering a specific species
profile is constant within regions, but different amongst them. These
groups also vary according to covariates, which in our case included
oceanographic, climatic, and physiographic parameters. \textbf{Results}:
Based on 142,646 seabirds recorded, from 80 species, all models
suggested two macro-scale assemblages (`northern' and `southern'),
except for the autumn presence-absence model that identified three
groups. The model results consistently show a biogeographic transition
at \textasciitilde34°S, near the latitude at which the East Australian
Current separates from the Australian continental slope. Sea surface
temperatures or sea surface salinities were selected in all final
models, further indicating a close relationship between seabird
assemblages and water masses. Results from both data types,
presence-absence and abundance, resulted in similar spatial and species
profile patterns. \textbf{Conclusion}: RCP models clearly identified two
seabird assemblages off the east coast of Australia, suggesting the
persistence of these groups at seasonal and macro spatial scales. Given
the rapid climatic changes that the East Australian Current is
experiencing and its importance in influencing distributions of
seabirds, the methods applied in our study could be replicated to assess
possible changes in seabird assemblages, and how they are affected by
changing environmental conditions.
\end{abstract}

\keywords{
    Assemblages
   \and
    Community ecology
   \and
    East Australian Current
   \and
    Mixture models
   \and
    Region of Common Profiles
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Bioregionalization is the discipline that identifies, delimits,
characterises, and names biogeographical regions (Morrone, 2018; Vilhena
\& Antonelli, 2015). Although there is often no agreement on the naming
systems among proposed regionalization schema (but see Ebach et al.,
2008), the hierarchical nature of bioregions is well recognized (Kreft
\& Jetz, 2010; Vilhena \& Antonelli, 2015). Thus, bioregions are usually
described within geographically larger classifications, such as
`provinces' or `realms' (Costello et al., 2017; Spalding et al., 2007).
The delimitation of bioregions is made through identifying sets of
species that are endemic or have high levels of co-occurrence, and are
distinct enough from other regions. In this sense, bioregions are
interpreted as a natural representations of common historical and/or
ecological processes shaped by physical and biological forces (Kreft \&
Jetz, 2010). In particular, the identification of bioregions allows
representation of the distribution of multiple species at a given time
and space, and relating these regions to environmental characteristics.
By doing so, bioregions can be interpreted as species assemblages (N. A.
Hill et al., 2017; Lyons et al., 2017). Given these features, bioregions
can be used as surrogates for biodiversity monitoring in data-poor
regions and in regional conservation planning (Ferrier, 2002; Spalding
et al., 2007), playing a fundamental role in conservation biogeography
(Whittaker et al., 2005).

Seabird assemblages have been widely described and related to
environmental characteristics (Hyrenbach et al., 2007; Ribic et al.,
1997; Woehler et al., 2003). Considering the hierarchical nature of
oceanographic processes and their resulting features (Haury et al.,
1978), it is not surprising that the distribution of highly mobile
seabirds are linked differently across temporal and spatial scales
(Haury et al., 1978; Hunt \& Schneider, 1987). In the marine
environment, large scale (\textgreater100 km length scale) oceanographic
features (e.g.~boundary currents and water masses) act as terrestrial
biomes (based on vegetation), and influence seabird assemblage
structure--i.e.~the occurrence of species and their relative numerical
composition (Commins et al., 2014; Hyrenbach et al., 2007; Ribic et al.,
1997). In contrast, local abundance is related to meso- (10--100 km) and
sub-mesoscale (1--10 km) processes that aggregate prey and therefore
predators, such as fronts and eddies (Scales, Miller, Hawkes, et al.,
2014; Weimerskirch, 2007). Such processes, although relatively
ephemeral, are primarily driven by physical and atmospheric forces, and
are thus somewhat predictable in space and time, providing consistent
feeding areas for marine predators when they occur (Davoren, 2013;
Hazen, Suryan, et al., 2013). Sea surface temperatures are a key driver
of seabird distributions, at individual- and assemblage-levels,
operating over a range of spatial and temporal scales from macro
(e.g.~water masses) to fine (e.g.~fronts) (Dunlop et al., 1988;
Quillfeldt et al., 2015; Ribic et al., 1997). Despite extensive surveys
(e.g. Ballance et al., 1997) and tracking studies (e.g. Carneiro et al.,
2020), many oceanic regions still lack information on the distributions
and abundances of seabird species, and how they assemble in space and
time.

One area in need of better understanding of seabird distributions is
Australia, considered a global seabird hotspot (Karpouzi et al., 2007;
Ramírez et al., 2017). Given the region's importance for seabird
diversity, there are surprisingly few studies on seabird distributions
at sea (Bernard et al., 2021; Mott \& Clarke, 2018). Off northwest
Australia, at-sea surveys, multivariate analyses (Dunlop et al., 1988)
and ensemble species distribution models of species assemblages (Lavers
et al., 2014) showed that habitat use by tropical seabird species was
related to sea surface temperatures and salinity. Dunlop et al. (1988)
suggested that salinity was the main variable influencing species'
distributions in oceanic waters while temperature was the key factor for
species on the continental shelf. In contrast, there are few studies at
large spatial and temporal scales off eastern Australia. T. A. Reid \&
Hindell (2000) and Evans et al. (2021) analysed seabird assemblages and
their relationships with oceanographic and prey data at coarse scales
(tens of km) off southeast Tasmania. The remaining studies are mostly
descriptive accounts locally (Blaber \& Milton, 1994; Norman, 1992), or
focused on a single species (Priddel et al., 2014).

The East Australian Current (EAC) is the major western boundary current
in the South Pacific Ocean, and the dominant boundary current off
eastern Australia (Ganachaud et al., 2014). The EAC has been extensively
studied, although research into the links between the EAC's oceanography
and marine predators are not well known. The dynamic features associated
with the EAC, such as mesoscale eddies and upwelling, are likely to
influence marine biota (Scales, Miller, Hawkes, et al., 2014; Suthers et
al., 2011). While most of the biological studies in this system were
focused on plankton entrapment within the cores of eddies (Firme et al.,
2023; Garcia et al., 2022; Hassler et al., 2011), there is a growing
body of work focusing on fisheries (Brieva et al., 2015; Hobday, 2010;
Young et al., 2011). These studies emphasized how the probability of
occurrence of a given taxon, and the structure of species assemblages,
were driven by the EAC. However, the relationships between oceanography
and marine megafauna within the EAC are poorly known (Suthers et al.,
2011). Some progress has been made using shark and ray movement data
(Jaine et al., 2014; Niella et al., 2020; Niella et al., 2022), and
satellite tracking has also demonstrated that increased sea surface
temperatures in the EAC decreases little penguin \emph{Eudyptula minor}
foraging success (Carroll et al., 2016), but information at the
assemblage-level for seabirds in the EAC is absent.

Marine species range shifts are more likely to occur faster than
terrestrial species due to their higher sensitivity to warming (Lenoir
et al., 2020; Pinsky et al., 2019). As a result, climate change is
rearranging the structure of species assemblages around the globe (Antão
et al., 2020; Dornelas et al., 2014; Poloczanska et al., 2013). Top
predators, for instance, may change their core habitats (Hazen,
Jorgensen, et al., 2013; Reisinger et al., 2022), with unprecedented
population consequences (Woehler \& Hobday, 2023). For seabirds, Péron
et al. (2010) and Sojitra et al. (2022) showed that sea surface
temperature anomalies affected the occurrence and abundances of
albatrosses and petrels species with latitude, thus affecting assemblage
structure. In eastern Australia, changing EAC conditions have affected a
range of marine species. For example, changes in zooplankton assemblage
structure and fish distribution patterns have been recorded and
attributed to warming water temperatures, and shifts in shark
distributions are expected (C. R. Johnson et al., 2011; Niella et al.,
2022). A severe marine heatwave driven by a poleward strengthening of
the EAC, showed drastic environmental impacts including disease
outbreaks in aquaculture farms and wildlife mortalities (Oliver et al.,
2017). Given that the EAC is strengthening and extending farther
southward and getting warmer (Cai et al., 2005; Phillips et al., 2022),
biological changes are expected to occur (Hobday, 2010; Poloczanska et
al., 2007).

In light of eastern Australia's importance for seabirds, and the recent
attention on how predicted climate change will influence the EAC's
marine biota (Poloczanska et al., 2007), an assessment of the
contemporary spatial distributions of seabirds within the EAC at large
temporal and spatial scales is timely. Seabirds can serve as sentinels
for tracking ecological changes in the EAC in many ways (e.g. Sydeman et
al., 2012). Surprisingly though, seabird distribution is
under-represented in climate change studies, despite being highlighted
as a priority topic (Grémillet \& Boulinier, 2009; Sydeman et al.,
2012).

In this study, we aim to identify assemblages of seabird species within
the EAC and adjacent areas, and relate these assemblages to
environmental conditions, by using at-sea surveys data and a
state-of-the-art modelling technique. We use the Region of Common
Profiles {[}RCP; Foster et al. (2013){]}, a mixture model regarded as
one of the best available tools for describing bioregions (N. A. Hill et
al., 2020; S. N. C. Woolley et al., 2020). RCP is a model-based
approach, and has a range of advantages over traditional algorithm-based
statistics (N. A. Hill et al., 2020; Warton et al., 2015). Eastern
Australia supports a high diversity of seabirds and offers wide
environmental gradients over a near 40° latitudinal range, from the
tropics to the cool temperate. We hypothesised to find different seabird
assemblages relating to different environmental conditions within the
EAC. In addition, presence-absence and abundance data were used to test
if the different numerical resolutions would result in the
identification of similar assemblages. This study provides a repeatable
framework on which future studies can compare to assess how future
changes in the EAC may influence the distribution of seabird
assemblages.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{study-area}{%
\subsection{Study area}\label{study-area}}

The EAC originates from the South Equatorial Current (SEC), a tropical
nutrient-poor current. When the SEC's North Caledonian Jet meets the
Queensland Plateau (\textasciitilde18°S), it splits into a north-flowing
branch towards the Gulf of Papua and a south-flowing branch, which marks
the start of the EAC (Ganachaud et al., 2014; Kessler \& Cravatte,
2013). The South Caledonian Jet completes the EAC system at
\textasciitilde22°S after encountering the shelf break in the southern
portion of the Great Barrier Reef (\textbf{Fig. 1a}; Kessler \& Cravatte
(2013); Ganachaud et al. (2014)). \textbf{SST/SSS characteristics? ROB?}
The EAC interacts with the bathymetry of the seafloor, giving rise to
oceanographic features and processes, such as smaller currents,
upwellings, and persistent dynamic eddies (Oke, Roughan, et al., 2019;
Oke \& Middleton, 2000; Ridgway \& Dunn, 2003). The EAC flow polewards
and is primarily associated with the shelf break, but at
\textasciitilde32°S it detaches from the shelf and spawns an area of
eddy generation that dissipates throughout the western Tasman Sea (Mata
et al., 2006; Oke, Roughan, et al., 2019), the `EAC eastern extension'
(Oke, Pilo, et al. (2019); often termed the `Tasman Front'). Although
most EAC energy is converted into eddies, the EAC still runs southwards
following the shelf break towards Tasmania (the `EAC southern
extension'; \textbf{Fig. 1a}), carrying eddy-like features as a
transient current (Oke, Roughan, et al., 2019).

\hypertarget{seabird-and-environmental-data-sources}{%
\subsection{Seabird and environmental data
sources}\label{seabird-and-environmental-data-sources}}

Seabird data were gathered aboard \emph{RV Investigator} and \emph{RV
Falkor} on 15 voyages, between 2016/17--2020/21, spanning over
\textasciitilde37° of latitude (\textbf{Fig. 1b}; \textbf{Table S1}).
One or more observers continuously recorded seabirds within a 300 m
forward quadrant (eye height 25 m a.s.l) from sunrise to sunset hours
while the ship was underway. The species, the numbers of birds, and
their behaviours were logged in real time into a ship-based web server.
\textbf{Environmental parameters for atmospheric and oceanographic
conditions were logged continuously by automated-ship based sensors,
with seabird observations supplementing relevant environmental
conditions such as visibility or associations with marine mammals.}
Individual records received a date/time/geographic coordinate label
(multi-species records have the same label but are logged as separate
entries for each species). Birds following the vessels were excluded
from analyses. Seabird observation protocols followed Raymond et al.
(2010), in use around Australia since 1980.

To investigate and predict species' distributions and relationships with
environmental variables, a considered choice of \textbf{broad-scale
climatic and oceanographic} covariates is important (Barbet-Massin \&
Jetz, 2014). We used three physiographic variables and seven
oceanographic variables, from which a further three climatic layers were
calculated (\textbf{Table 1}). These variables are the most frequently
used in seabird-environment relationship studies (Tremblay et al.,
2009), capturing processes at a range of spatial and temporal scales
from macro-scale water-mass properties to prey availability in meso- to
submesoscale oceanographic features (Haury et al., 1978; Scales, Miller,
Hawkes, et al., 2014; Weimerskirch, 2007).

\hypertarget{data-treatment}{%
\subsection{Data treatment}\label{data-treatment}}

Data wrangling, modelling and visualization were done in R 4.2.0 (R Core
Team, 2022); the full list of packages can be found in the Supplementary
Material \textbf{XX}. Code and the modelling dataset (see
\ref{sec:modellingdatasets}) are available in \textbf{Daudt et al.~2023
(Zenodo repo)}.

\hypertarget{environmental-layers}{%
\subsubsection{Environmental layers}\label{environmental-layers}}

The spatial and temporal persistence of oceanographic features such as
eddy kinetic energy (`eke') and thermal fronts (sea surface temperature
gradient: `sst\_grad') may influence the movements and distributions of
seabirds by offering a higher likelihood of finding food (Scales,
Miller, Embling, et al., 2014; Weimerskirch, 2007). Thus, `eke' and
`sst\_grad' environmental layers were calculated before extracting
environmental data for each seabird record. For both `eke' and
`sst\_grad', we used their original cell spatial resolution
(\textbf{Table 1}) and calculations were done layer-by-layer to match
seabird data on the finest spatial scale possible. Eddy kinetic energy
was calculated as \(0.5 (u'² + v'²)\), where \(u' = u - \bar{u}\) and
\(v' = v - \bar{v}\) are the surface zonal (latitudinal) and meridian
(longitudinal) geostrophic currents, respectively, using a rolling mean
of 90 days as \(\bar{u}\) and \(\bar{v}\) (Sun et al., 2022). The
spatial `sst\_grad' for each cell was calculated by averaging the
maximum value for each neighbouring cell (average neighbourhood
maximum), from north-to-south (NS) and west-to-east (WE); then, the
magnitude of the gradient was defined as \(NS² + WE²\) (Burrows et al.,
2011). In addition, we calculated climatic seasonal average layers for
`eke' and `sst\_grad' based on the 11-year data for 2011--2021 as a
proxy for persistent oceanographic features, and the `eke' standard
deviation over the same period (named with the prefix `clim\_' in
\textbf{Table 1}). Seabed slope was calculated using `raster::terrain()'
(Hijmans, 2022a), and distance from the coast was calculated through
`sf::st\_distance()' (Pebesma, 2018) using Australia and Tasmania as the
mainland (i.e.~ignoring smaller islands). Then, for each seabird record,
the most proximate value for each environmental layer (original or
calculated) was extracted using built-in functions from `raster' 3.5-21,
`terra' 1.6-7, and `sf' 1.0-8 packages (Hijmans, 2022a, 2022b; Pebesma,
2018).

\hypertarget{modellingdatasets}{%
\subsubsection{Data}\label{modellingdatasets}}

\label{sec:modellingdatasets}

Seabirds are highly mobile animals, and many species migrate and/or
disperse during their non-breeding periods (Schreiber \& Burger, 2002).
Therefore, seasonal differences in species composition were expected
(e.g. K. Reid et al., 2023), and in the probabilities of occurrence and
in total numbers. Thus, we built separate modelling dataset for each
season. We aggregated seabird data by 1° latitude \(\times\) 1°
longitude grid cells (c.~110 \(\times\) 110 km) and season (\textbf{Fig.
1b}), by summing seabirds from the same species. To test if the
presence-absence and abundance (raw counts) data would result in similar
assemblage patterns, we built two species matrices, where the
presence-absence matrix transformed values \textgreater1 to 1. To avoid
numerical instabilities in the models prompted by unusual or rare
species, we kept only species occurring in at least six grids within
each season (summer = 12 species; autumn = 21; winter = 20; spring =
33). The associated environmental data of each seabird record were
averaged within the same grid cells and seasons, and environmental data
were standardised before modelling.

\hypertarget{statistical-modelling}{%
\subsection{Statistical modelling}\label{statistical-modelling}}

We used the mixture model Region of Common Profiles (RCP) to identify
and describe seabird species assemblages. RCP is a one-stage
bioregionalization approach, as it identifies bioregions by specifying a
statistical model that simultaneously relates response (multi-species
data) and predictor (environmental data) (N. A. Hill et al., 2020; S. N.
C. Woolley et al., 2020). Given a set of environmental variables, the
model can predict assemblage probabilities of sites where no biotic
(observational) data exist. As a model-based approach, RCPs can be
compared by information criteria, present uncertainty around the
results, and check for model assumptions through assessing distribution
of residuals (N. A. Hill et al., 2020; Warton et al., 2015).

The model is specified in terms of assemblages and site-specific
probabilities. There are \(K\) assemblages (i.e.~RCPs) that are common
across the study area, \(K = (1, ..., k)\). The outcome variable is a
vector \(y_{i} = (y_{i1}, ... y_{iS})\), that give the count or
presence-absence of each of the \(S\) species at site \(i\). Depending
on whether we model the counts or presence-absence data, the assemblages
are defined in terms of the expected frequency or relative frequency of
the species, respectively. An output of the model is the estimated
probability that site \(i\) has each of the \(K\) possible assemblages.
RCPs are a multivariate adaptation of a mixture-of-experts model; as
such, the mixing proportions (assemblage structure) vary according to
covariates (predictors) (Foster et al., 2013). However, the final
estimated assemblage structure is assumed to be constant for all sites
belonging to the \(K\) assemblage (Foster et al., 2013). As such, RCPs
can be directly interpreted as assemblages, and their species profiles
as the assemblage structure.

Before fitting models, covariates were checked for collinearity within
each seasonal dataset. If a pairwise collinearity of
\(\left|x\right| >0.7\) was found we removed one of the variables
(Dormann et al., 2013), favouring direct measures over the `climatic'
layers we calculated (e.g.~if `sst\_grad' and `clim\_sst\_grad' were
correlated, we chose `sst\_grad'). Thus, `full models' differed between
seasons as the retained covariates varied, although most covariates were
common for all seasons.

Due to seabird migration and dispersion, seasons could act as `sampling
artifacts' (Foster et al., 2017), thus biasing results (Ponti \&
Sannolo, 2023). In light of the difference among seasons in the numbers
of species that met our occurrence threshold (see
\ref{sec:modellingdatasets}), we decided to use RCP models by season and
data type (presence-absence and abundance). RCP models were used to
relate the response variables (i.e.~presence-absence or abundance of
each seabird species in each grid cell) to the independent predictor
covariates (i.e.~averaged values of environmental variables in each
cell). Presence-absence data were modelled based on a Bernoulli model,
whereas abundance data based on a Negative Binomial model. RCPs were
fitted using the `ecomix' 1.0.0 package (S. Woolley et al., 2022), with
the following workflow: (i) find the best number of groups (\(K\)) that
describes the data, (ii) re-fit the model by fixing \(K\) and selecting
optimal combination of covariates, (iii) check residuals, (iv) bootstrap
the best model to estimate uncertainty around parameter estimates, and
(v) produce results.

To identify the best number of RCP groups (\(K\)) that described the
data, we used full models (all covariates) with 100 random start numbers
(see Foster et al., 2013) between \(k = (2, ..., 7)\) groups, and chose
the number of RCPs that minimized the Bayesian Information Criterion
(BIC). Then, we fixed \(K\) at the optimal number and fitted models for
every possible linear, additive combination of covariates. We chose the
best model as the one with the lowest BIC. Visual inspection of
randomized quantile residual plots modified for mixture models was used
to verify model assumptions (Dunstan et al., 2013). To obtain estimates
of uncertainty for model parameters (confidence intervals), we used the
default settings of `ecomix::regional\_mix.bootstrap()', which applies a
Bayesian bootstrap 1000 times. A custom code was written to generate
partial plots for exploring the influence of covariates on RCP
membership.

We built a seasonal 11-year (2011--2021) average layer for predicting
model results for each environmental covariate. These variables were
then averaged within 1° latitude \(\times\) 1° longitude grid cells, by
season, between 9--47°S and 141--161°E. Note that `clim\_eke\_mean' and
`clim\_sst\_grad' were processed beforehand as they may characterise
persistent oceanographic features (described above) and thus represent
the same value as `eke' and `sst\_grad'. Data for prediction were
standardised using the same mean and standard deviation from the
modelling dataset.

Finally, we ran rarefaction curves based on the presence-absence matrix
and the assigned RCP group for each sample (grid cell) to assess if the
amount of information in the RCP model was representative. Rarefaction
curves were fit using the `iNEXT' 3.0.0 package (Chao et al., 2014;
Hsieh et al., 2016), and results are presented based on alpha diversity
(i.e.~species richness).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{summary-of-species-richness-and-numbers}{%
\subsection{Summary of species richness and
numbers}\label{summary-of-species-richness-and-numbers}}

A total of 10,261 seabird records was obtained during the 15 voyages,
comprising 142,646 individuals from 80 species, belonging to 11 families
and five orders. The five most abundant species made up 83\% of total
birds counted (43.0\% short-tailed shearwater \emph{Ardenna
tenuirostris}, 17.2\% fairy prion \emph{Pachyptila turtur}, 14.3\%
wedge-tailed shearwater \emph{Ardenna pacifica}, 5.2\% sooty tern
\emph{Onychoprion fuscatus}, 3.2\% shy albatross \emph{Thalassarche
cauta}). The five most frequently encountered species were \emph{A.
pacifica} (frequency of occurrence = 35.3\%), \emph{T. cauta} (33.2\%),
brown booby \emph{Sula leucogaster} (25.4\%), red-footed booby \emph{S.
sula} (25.4\%) and black-browed albatross \emph{T. melanophris} (25\%)
(\textbf{Fig. S1}). An increase in species richness towards the south of
the study region was clear in all but the winter season (\textbf{Fig
S2}). Noticeably higher counts were made around Tasmania, possibly due
to the `colony effect' and/or the presence of Southern Ocean species
(\textbf{Fig. S2}). However, one cell in the Coral Sea had counts of the
same order of magnitude as off Tasmania in summer.

\hypertarget{rcp-models}{%
\subsection{RCP models}\label{rcp-models}}

For all models, two RCPs was the optimal number of seabird groups,
except for the presence-absence-autumn model, which indicated three RCPs
as the best fit (\textbf{Fig. S3}). The best fitting models generally
had 1 or 2 covariates selected, although the abundance-autumn model had
five covariates (\textbf{Table 2}). The best models showed a good fit
with no apparent deviations from the assumptions (\textbf{Fig. S4}). All
the best models selected `sst' or `sss', as essential for characterising
water masses. The probability of belonging to one or another RCP varied
its `sst' threshold between 18--21 °C, or `sss' 35.1 PSU for summer
(\textbf{Fig. 2}, \textbf{Fig. S5}), similar to the EAC's temperature
and salinity thresholds. Furthermore, most of the best models selected
`bat', suggesting a gradient between `coastal' and `oceanic' taxa based
on bathymetry (\textbf{Fig. S5}). Of particular note, proxies for
persistent oceanographic features that may aggregate food resources (the
`climatic' variables) for seabirds were selected only in the autumn and
spring abundance models.

Grid point-predictions are the assignment of the RCP with the highest
probability value for a given cell. When predicting RCP onto space,
patterns of point-predictions suggest a biogeographic transition in
seabird species around \textasciitilde34°S, with both presence-absence
and abundance models pointing to similar results (\textbf{Fig. 3}).
These patterns are further reinforced when looking at the mean
probability along with the lower and upper confidence intervals of
predictions (\textbf{Fig. 4}, \textbf{Fig. S6}), showing that around
\textasciitilde34°S, the `northern' RCP tended to have higher
probabilities in coastal regions, whereas the `southern' RCP spread
northward offshore. The presence-absence-autumn model (three RCPs) had a
similar broad-scale pattern, although the southern RCP is farther south,
and a strong `Tasmania/Bass Straight' assemblage appears (\textbf{Fig.
3}, \textbf{Fig. S6}).

The broad pattern of two large-scale assemblages is reflected in their
species compositions. Although there are overlaps among the seabird
groups, both presence-absence and abundance models captured typical
tropical species in the northern RCP (e.g.~boobies \emph{Sula} spp.,
lesser frigatebird \emph{Fregata ariel}) and typical Southern Ocean
species in the southern RCP (e.g.~small albatrosses \emph{Thalassarche}
spp., fairy prion \emph{P. turtur}) (\textbf{Fig. 5}, \textbf{Fig. S7}).
The presence-absence-autumn model, which had three RCPs, also showed the
characteristic taxa representing the `northern' and `southern' RCPs.
However, the intermediate `Tasmania/Bass Straight' assemblage showed a
mixture of both groups with shy albatross and Australasian gannet
\emph{Morus serrator} prominent with high probabilities of occurrence.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

This study showed that seabird assemblages off eastern Australia were
correlated with oceanographic conditions, and that the East Australian
Current appeared to be key in structuring seabird distributions at meso-
to macro scales (\emph{sensu} Hunt \& Schneider, 1987). Our results are
based on a multivariate statistical model (N. A. Hill et al., 2020),
which brings advantages over classical methods for describing species
assemblages. RCP models allow estimating uncertainty around the point
probability of a sample (grid cell) belonging to an assemblage, thus
representing the fuzziness between bioregion boundaries. These
transition zones between bioregions are essential for monitoring
biological changes due to environmental changes (Morrone, 2023). As
such, the boundary between the two consistently identified bioregions in
this study may be a crucial region to monitor in future, given
noticeable anomalies in the EAC at this same region (Phillips et al.,
2022). Our results draw an important baseline for seabird assemblages in
a world hotspot for seabirds.

\hypertarget{seabird-bioregions-off-eastern-australia}{%
\subsection{Seabird bioregions off eastern
Australia}\label{seabird-bioregions-off-eastern-australia}}

The relationships between seabird assemblages and water masses has been
known for a long time (Elphick \& Hunt, 1993; Pocklington, 1979; Ribic
et al., 1997). Thus, it is not surprising that two seabird assemblages
were identified off eastern Australia. However, the spatial consistency
of these patterns throughout seasons is surprising and remarkable.
Resident species seem to characterise these assemblages. Shy albatross
(Hedd et al., 2001; Mason et al., 2018) and fairy prions (Fromant et
al., 2022) in the southern assemblage, and boobies (Bunce, 2015; Miller
et al., 2018) and wedge-tailed shearwaters (McDuie et al., 2015;
Weimerskirch et al., 2020) in the northern assemblage, were always
profiled with higher probabilities of occurrence and predicted
abundances compared to other species in the same groups. Given that
changes in assemblages over time are typically driven by a few species
(Gotelli et al., 2022), and that those species are likely residents
(Souza \& Santos, 2023), the relationships between these resident
seabirds and the EAC system over the broad latitudinal range detailed
here should be closely monitored. Elsewhere, coastal seabirds such as
little penguin are signalling the adverse effects that climate change
might have in changing seabird behaviours and decreasing foraging
success (Carroll et al., 2016; Woehler \& Hobday, 2023), which
ultimately influence their breeding success and population dynamics (B.
Johnson \& Colombelli-Négrel, 2021). Thus, the species profiles and the
spatial predictions of assemblages from RCP models could be used as
tools for monitoring future seabird responses to the fast-changing EAC.

The seasonal models also captured the influence of migratory species,
particularly in the southern assemblage. The summer and autumn models
reflected the high abundance of short-tailed shearwater before they
depart on their trans-equatorial migration (Carey et al., 2014); the
presence of the Australasian gannet was picked up by the autumn and
winter models (Ismar et al., 2011); and the white-headed petrel
\emph{Pterodroma lessonii} movements during chick-rearing and pre-laying
exodus (Taylor et al., 2021) were detected in autumn and spring models,
respectively. Given the geographic location of the southern assemblage,
however, we believe sub-Antarctic taxa were not well represented in our
models due to their low number of occurrences and relative paucity of
survey effort to date (\textbf{Fig. S1}; hence they were removed before
modelling). Additional surveys may reveal the seasonal northward pulse
of these taxa into the southern assemblage (e.g.~albatrosses
\emph{Diomedea} spp. and \emph{Phoebetria} spp.), given that more effort
translates as higher likelihood of recording these species. In turn,
including these taxa in the models could result in a more stable `three
assemblage pattern' (as revealed by the autumn presence-absence model)
in all but the summer season, when the majority of Antarctic species are
breeding (S. Marchant \& Higgins, 1990).

Our analyses clearly revealed a `northern' and a `southern' assemblage
of seabirds off eastern Australia within the EAC. There is a noticeable
link with the EAC dynamics: the boundaries of RCPs occurred at
\textasciitilde34°S, near the average separation latitude
(\textasciitilde32°S) of the EAC and its eastward extension into the
Tasman Sea (Oke, Roughan, et al., 2019). However, the
`northern/southern' assemblage pattern identified herein does not agree
with previously-proposed large-scale regionalizations, such as the
pelagic Longhurst Provinces (Longhurst, 2007), the Marine Ecoregions of
the World {[}MEOW; Spalding et al. (2007){]}, or the Large Marine
Ecosystems {[}LME; Sherman (1994){]}. The MEOW schema was based on
coastal and shelf biota and was meant to represent habitats to the 200 m
isobath, which clearly does not encompass the entirety of seabird
habitats. The LME schema is characterised by physiographical,
hydrological and productivity data but has a strong focus on practical
management of areas and thus there is a political component embedded
within it, which does not align with seabird distributions. Longhurst
(2007) designated only one biogeochemical province for eastern Australia
based on productivity patterns, which may work well at larger (macro)
scales, but appears to be oversimplified for the EAC.

The results do agree with the marine biogeographic units proposed by
Costello et al. (2017), who identified a boundary at \textasciitilde34°S
between their `Coral Sea' and `South Australia' realms. Their approach
used 65,000 taxa, from benthic/pelagic and coastal/offshore habitats,
thus integrating more complexity of the biogeographic boundaries in the
marine environment. Other biological data also suggest
`northern/southern' assemblages off eastern Australia, including tuna
and billfish {[}also highly mobile top predators; Reygondeau et al.
(2012){]}, mesopelagic fauna {[}which includes seabird prey; Sutton et
al. (2017){]} and deep-sea invertebrates and fish (O'Hara, Williams,
Althaus, et al., 2020; O'Hara, Williams, Woolley, et al., 2020). All
these studies focused mainly on describing biological assemblages, as
opposed to classifying environments. It is worth noting that the
Longhurst Provinces approach was updated to include SST, SSS and
bathymetry (the same variables retained by our best RCP models) and a
temporal component to the classification (Reygondeau et al., 2013). The
updated provinces are more similar to the seabird assemblages described
here, although still not a perfect match, which could be partially
explained because the seabird assemblage at EAC region includes several
long-distance migratory species. Australia's bioregionalization has a
long story of disagreements (Ebach, 2012), however, the inherent link
that RCPs develop between environment and biological data can be a
strong tool to support marine bioregionalization.

\hypertarget{fitting-rcp-models-to-high-mobile-animals}{%
\subsection{Fitting RCP models to high mobile
animals}\label{fitting-rcp-models-to-high-mobile-animals}}

To the best of our knowledge, this is the first study applying RCPs to
highly mobile animals such as seabirds. RCP models have been previously
applied to plants (Lyons et al., 2017), and relatively sedentary
animals, such as macrobenthic organisms (Pantiukhin et al., 2021) and
demersal fish (N. A. Hill et al., 2017). Oceanic seabirds can have daily
movements of hundreds of kilometres (e.g. McDuie et al., 2015), during
which they traverse a variety of habitats and environmental gradients.
As such, capturing the distribution of highly mobile species, and their
relationships with environmental parameters is not a trivial task, and
much harder than for sessile organisms.

While remote tracking studies are advancing our understanding on how
individual seabirds use their environment at fine scales (Tremblay et
al., 2009), they fail to perceive how species assemblages interact with
the environment and each other. Therefore, using at-sea counts of
seabirds is a well-established method to capture these relationships
(Ainley et al., 2012; Hyrenbach et al., 2007; Woehler et al., 2003).
Further, the spatial scales on which assemblages relate to their
environment is variable and species-specific (Hunt \& Schneider, 1987;
Ribic et al., 1997; Scales et al., 2017). Understanding the biology of
the studied group is thus essential to set the appropriate scales of
analyses. Our choice of 1° latitude \(\times\) 1° longitude was mainly
determined by data availability reflecting survey effort, and the
trade-off between the number of species per grid/season (see also
\ref{sec:limitations}). Nonetheless, the chosen spatial scale is similar
to other seabird distribution analyses conducted elsewhere (Harris et
al., 2007; Santora \& Veit, 2013; Sojitra et al., 2022). Therefore, for
describing seabird assemblages at a macro-scale, the chosen resolution
seems to had been a good fit, accords with studies elsewhere, and the
resultant models described patterns of seabird species' composition
consistent with our current understanding of species' ecologies and life
history strategies.

RCP models based on presence-absence and abundance data are able to
complement each other. Previous studies using RCPs have opted to use
only one type of response data. Lyons et al. (2017), N. A. Hill et al.
(2017), and Pantiukhin et al. (2021) all used presence-absence data,
whereas Lee et al. (2019) and Receveur et al. (2020) used abundance
(count) data. The added information contained in abundance data make
them inherently more powerful than binary presence-absence data.
However, the use of different data inputs in multivariate, biodiversity
assessments and assemblage descriptions has shown contrasting patterns
at different spatial scales. At larger spatial and temporal scales,
presence-absence data describe similar patterns to abundance data
(Anderson et al., 2005), as we found. In contrast, at smaller scales,
abundance data may highlight local conditions and therefore reveal
hidden patterns at coarse to fine scales (Anderson et al., 2005; R.
Marchant, 1990; Melo, 2005; Waters, 2008). If combined with ecological
knowledge of a particular group, the information contained in both
presence-absence and abundance data could be leveraged to better
interpret the resulting patterns. For example, seabirds will likely
occur in specific water masses or currents (Hunt \& Schneider, 1987;
Ribic et al., 1997), whereas their abundance is related to coarse- to
fine-scale processes that enhance prey availability (Hunt \& Schneider,
1987; Weimerskirch, 2007). Future studies assessing bioregions of highly
mobile animals could describe the compositions of assemblages using
smaller resolution grid cells and presence-absence models. Then,
abundance models using higher resolution grid cells could improve our
understanding of how these animals relate to their environment and
possibly reveal multi-species aggregation areas (therefore `new
bioregions') likely nested within the bioregions described at large
scales. Such a hierarchical approach would require substantial
concurrent observational and environmental data sets, which we suspect
not exist at present.

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\label{sec:limitations}

Although the results suggest consistent patterns, there are a number of
caveats to our analyses. Seabird data were gathered during
multi-disciplinary cruises, whose sampling designs were not specifically
planned to account for spatial ecology of seabirds. As a result, there
are spatial discontinuities in the data, which might have limited the
ability of our models to detect the relationships between the seabird
assemblages and the full spectrum of environmental predictors. Even
though Australia is a hotspot for seabirds (Karpouzi et al., 2007;
Ramírez et al., 2017), there is a paucity of information on seabird
distributions in the region (Bernard et al., 2021; Mott \& Clarke,
2018). Dedicated seabird surveys would help to advance seabird research
in this region, and improve our understanding of the patterns described
here. By increasing the amount of data, future studies could set the
grid cells at higher resolution, and potentially reveal assemblages at
finer spatial and temporal scales. In addition, ideally, we would have
had more information within each grid cell by season (our sampling
units). Nonetheless, rarefaction curves show that the sample coverage
(see Chao \& Jost, 2012) was high at the RCP level, although there is
room for detecting more species in most of them (\textbf{Fig S8}).

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

We applied a robust, repeatable method to assess seabird assemblages off
eastern Australia within the EAC for the first time. The models point
toward a consistent pattern of a `northern' and a `southern' seabird
assemblage, where the spatial division is attributable to the eastward
separation of the EAC from the coast, and that SST seems to be the key
variable discriminating these groups. The EAC has penetrated poleward
over the past half-century (K. L. Hill et al., 2008; Li et al., 2022), a
trend which is anticipated to continue into the future (Yang et al.,
2016), and extreme weather events such as marine heatwaves to happen
more often (Oliver et al., 2018). As such, seabirds will have to respond
to these changes (Woehler \& Hobday, 2023). Southern Ocean albatrosses
off southeast Australia, for instance, have shown `early stages of a
southward range shift' due to environmental variability (Sojitra et al.,
2022). Resident species such as the little penguins and wedge-tailed
shearwater change their behaviour and have decreased foraging success
when SST is higher (Carroll et al., 2016; Peck et al., 2004). Given that
eastern Australia is a biodiversity and numerical hotspot for seabirds,
it is imperative to establish a monitoring program allowing the
assemblages and species' profiles presented here to be used as the
baseline for understanding possible effects of a warming ocean on
seabird distribution. In particular, the transition zone between
assemblages (Morrone, 2023), associated with the eastern extension of
the EAC, may be fundamental to effectively monitoring strengthening of
the EAC and the subsequent impact on marine biota.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We thank all observers who assisted in data collection, the crews of RV
\emph{Investigator} and RV \emph{Falkor}, and the Australasian Seabird
Group (BirdLife Australia) for logistical support. The authors would
like to thank Dr.~Skipton N.C. Woolley, \texttt{\{ecomix\}} developer,
who kindly helped NWD build preliminary models and gave advice in
several stages of model construction. NWD is supported by a Doctoral
scholarship from the University of Otago.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-ainley2012}{}}%
Ainley, D., Ribic, C., \& Woehler, E. (2012). Adding the ocean to the
study of seabirds: a brief history of at-sea seabird research.
\emph{Marine Ecology Progress Series}, \emph{451}, 231--243.
\url{https://doi.org/10.3354/meps09524}

\leavevmode\vadjust pre{\hypertarget{ref-anderson2005}{}}%
Anderson, M. J., Connell, S. D., Gillanders, B. M., Diebel, C. E., Blom,
W. M., Saunders, J. E., \& Landers, T. J. (2005). Relationships between
taxonomic resolution and spatial scales of multivariate variation.
\emph{Journal of Animal Ecology}, \emph{74}(4), 636--646.
\url{https://doi.org/10.1111/j.1365-2656.2005.00959.x}

\leavevmode\vadjust pre{\hypertarget{ref-antuxe3o2020}{}}%
Antão, L. H., Bates, A. E., Blowes, S. A., Waldock, C., Supp, S. R.,
Magurran, A. E., Dornelas, M., \& Schipper, A. M. (2020).
Temperature-related biodiversity change across temperate marine and
terrestrial systems. \emph{Nature Ecology \& Evolution}, \emph{4}(7),
927--933. \url{https://doi.org/10.1038/s41559-020-1185-7}

\leavevmode\vadjust pre{\hypertarget{ref-ballance1997}{}}%
Ballance, L. T., Pitman, R. L., \& Reilly, S. B. (1997). Seabird
community structure along a productivity gradient: importance of
competition and energetic constraint. \emph{Ecology}, \emph{78}(5),
1502--1518.
\url{https://doi.org/10.1890/0012-9658(1997)078\%5B1502:SCSAAP\%5D2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-barbet-massin2014}{}}%
Barbet-Massin, M., \& Jetz, W. (2014). A 40-year, continent-wide,
multispecies assessment of relevant climate predictors for species
distribution modelling. \emph{Diversity and Distributions},
\emph{20}(11), 1285--1295. \url{https://doi.org/10.1111/ddi.12229}

\leavevmode\vadjust pre{\hypertarget{ref-bernard2021}{}}%
Bernard, A., Rodrigues, A. S. L., Cazalis, V., \& Grémillet, D. (2021).
Toward a global strategy for seabird tracking. \emph{Conservation
Letters}, \emph{14}(3), e12804. \url{https://doi.org/10.1111/conl.12804}

\leavevmode\vadjust pre{\hypertarget{ref-blaber1994distribution}{}}%
Blaber, S., \& Milton, D. (1994). Distribution of seabirds at sea in the
{G}ulf of {C}arpentaria, {A}ustralia. \emph{Marine and Freshwater
Research}, \emph{45}(3), 445--454.

\leavevmode\vadjust pre{\hypertarget{ref-brieva2015}{}}%
Brieva, D., Ribbe, J., \& Lemckert, C. (2015). Is the East Australian
Current causing a marine ecological hot-spot and an important fisheries
near Fraser Island, Australia? \emph{Estuarine, Coastal and Shelf
Science}, \emph{153}, 121--134.
\url{https://doi.org/10.1016/j.ecss.2014.12.012}

\leavevmode\vadjust pre{\hypertarget{ref-bunce2015}{}}%
Bunce, A. (2015). Foraging behaviour of a declining population of brown
boobies ({\emph{Sula leucogaster}}) breeding in the Swain Reefs, Great
Barrier Reef. \emph{Emu - Austral Ornithology}, \emph{115}(4), 368--372.
\url{https://doi.org/10.1071/MU14037}

\leavevmode\vadjust pre{\hypertarget{ref-burrows2011}{}}%
Burrows, M. T., Schoeman, D. S., Buckley, L. B., Moore, P., Poloczanska,
E. S., Brander, K. M., Brown, C., Bruno, J. F., Duarte, C. M., Halpern,
B. S., Holding, J., Kappel, C. V., Kiessling, W., O'Connor, M. I.,
Pandolfi, J. M., Parmesan, C., Schwing, F. B., Sydeman, W. J., \&
Richardson, A. J. (2011). The pace of shifting climate in marine and
terrestrial ecosystems. \emph{Science}, \emph{334}(6056), 652--655.
\url{https://doi.org/10.1126/science.1210288}

\leavevmode\vadjust pre{\hypertarget{ref-cai2005}{}}%
Cai, W., Shi, G., Cowan, T., Bi, D., \& Ribbe, J. (2005). The response
of the Southern Annular Mode, the East Australian Current, and the
southern mid-latitude ocean circulation to global warming.
\emph{Geophysical Research Letters}, \emph{32}(23), L23706.
\url{https://doi.org/10.1029/2005GL024701}

\leavevmode\vadjust pre{\hypertarget{ref-carey2014}{}}%
Carey, M. J., Phillips, R. A., Silk, J. R. D., \& Shaffer, S. A. (2014).
Trans-equatorial migration of short-tailed shearwaters revealed by
geolocators. \emph{Emu - Austral Ornithology}, \emph{114}(4), 352--359.
\url{https://doi.org/10.1071/MU13115}

\leavevmode\vadjust pre{\hypertarget{ref-carneiro2020}{}}%
Carneiro, A. P. B., Pearmain, E. J., Oppel, S., Clay, T. A., Phillips,
R. A., Bonnet-Lebrun, A.-S., Wanless, R. M., Abraham, E., Richard, Y.,
Rice, J., Handley, J., Davies, T. E., Dilley, B. J., Ryan, P. G., Small,
C., Arata, J., Arnould, J. P. Y., Bell, E., Bugoni, L., \ldots{} Dias,
M. P. (2020). A framework for mapping the distribution of seabirds by
integrating tracking, demography and phenology. \emph{Journal of Applied
Ecology}, \emph{57}(3), 514--525.
\url{https://doi.org/10.1111/1365-2664.13568}

\leavevmode\vadjust pre{\hypertarget{ref-carroll2016}{}}%
Carroll, G., Everett, J. D., Harcourt, R., Slip, D., \& Jonsen, I.
(2016). High sea surface temperatures driven by a strengthening current
reduce foraging success by penguins. \emph{Scientific Reports},
\emph{6}(1), 22236. \url{https://doi.org/10.1038/srep22236}

\leavevmode\vadjust pre{\hypertarget{ref-chao2014}{}}%
Chao, A., Chiu, C.-H., \& Jost, L. (2014). Unifying species diversity,
phylogenetic diversity, functional diversity, and related similarity and
differentiation measures through Hill numbers. \emph{Annual Review of
Ecology, Evolution, and Systematics}, \emph{45}(1), 297--324.
\url{https://doi.org/10.1146/annurev-ecolsys-120213-091540}

\leavevmode\vadjust pre{\hypertarget{ref-chao2012}{}}%
Chao, A., \& Jost, L. (2012). Coverage-based rarefaction and
extrapolation: standardizing samples by completeness rather than size.
\emph{Ecology}, \emph{93}(12), 2533--2547.
\url{https://doi.org/10.1890/11-1952.1}

\leavevmode\vadjust pre{\hypertarget{ref-commins2014}{}}%
Commins, M. L., Ansorge, I., \& Ryan, P. G. (2014). Multi-scale factors
influencing seabird assemblages in the African sector of the Southern
Ocean. \emph{Antarctic Science}, \emph{26}(1), 38--48.
\url{https://doi.org/10.1017/S0954102013000138}

\leavevmode\vadjust pre{\hypertarget{ref-costello2017}{}}%
Costello, M. J., Tsai, P., Wong, P. S., Cheung, A. K. L., Basher, Z., \&
Chaudhary, C. (2017). Marine biogeographic realms and species
endemicity. \emph{Nature Communications}, \emph{8}(1), 1057.
\url{https://doi.org/10.1038/s41467-017-01121-2}

\leavevmode\vadjust pre{\hypertarget{ref-davoren2013}{}}%
Davoren, G. K. (2013). Distribution of marine predator hotspots
explained by persistent areas of prey. \emph{Marine Biology},
\emph{160}(12), 3043--3058.
\url{https://doi.org/10.1007/s00227-013-2294-5}

\leavevmode\vadjust pre{\hypertarget{ref-dormann2013}{}}%
Dormann, C. F., Elith, J., Bacher, S., Buchmann, C., Carl, G., Carré,
G., Marquéz, J. R. G., Gruber, B., Lafourcade, B., Leitão, P. J.,
Münkemüller, T., McClean, C., Osborne, P. E., Reineking, B., Schröder,
B., Skidmore, A. K., Zurell, D., \& Lautenbach, S. (2013). Collinearity:
a review of methods to deal with it and a simulation study evaluating
their performance. \emph{Ecography}, \emph{36}(1), 27--46.
\url{https://doi.org/10.1111/j.1600-0587.2012.07348.x}

\leavevmode\vadjust pre{\hypertarget{ref-dornelas2014}{}}%
Dornelas, M., Gotelli, N. J., McGill, B., Shimadzu, H., Moyes, F.,
Sievers, C., \& Magurran, A. E. (2014). Assemblage time series reveal
biodiversity change but not systematic loss. \emph{Science},
\emph{344}(6181), 296--299.
\url{https://doi.org/10.1126/science.1248484}

\leavevmode\vadjust pre{\hypertarget{ref-dunlop1988}{}}%
Dunlop, J., Wooller, R., \& Cheshire, N. (1988). Distribution and
abundance of marine birds in the Eastern Indian Ocean. \emph{Marine and
Freshwater Research}, \emph{39}(5), 661.
\url{https://doi.org/10.1071/MF9880661}

\leavevmode\vadjust pre{\hypertarget{ref-dunstan2013}{}}%
Dunstan, P. K., Foster, S. D., Hui, F. K. C., \& Warton, D. I. (2013).
Finite mixture of regression modeling for high-dimensional count and
biomass data in ecology. \emph{Journal of Agricultural, Biological, and
Environmental Statistics}, \emph{18}(3), 357--375.
\url{https://doi.org/10.1007/s13253-013-0146-x}

\leavevmode\vadjust pre{\hypertarget{ref-ebach2012history}{}}%
Ebach, M. C. (2012). A history of bioregionalisation in {A}ustralia.
\emph{Zootaxa}, \emph{3392}, 1--34.
\url{https://doi.org/10.11646/ZOOTAXA.3392.1.1}

\leavevmode\vadjust pre{\hypertarget{ref-ebach2008}{}}%
Ebach, M. C., Morrone, J. J., Parenti, L. R., \& Viloria, Á. L. (2008).
International Code of Area Nomenclature. \emph{Journal of Biogeography},
\emph{35}(7), 1153--1157.
\url{https://doi.org/10.1111/j.1365-2699.2008.01920.x}

\leavevmode\vadjust pre{\hypertarget{ref-elphick1993}{}}%
Elphick, C. S., \& Hunt, G. L. (1993). Variations in the distributions
of marine birds with water mass in the northern Bering Sea.
\emph{Condor}, \emph{95}(1), 33--44.
\url{https://doi.org/10.2307/1369384}

\leavevmode\vadjust pre{\hypertarget{ref-evans2021a}{}}%
Evans, R., Lea, M.-A., \& Hindell, M. A. (2021). Predicting the
distribution of foraging seabirds during a period of heightened
environmental variability. \emph{Ecological Applications}, \emph{31}(5),
e02343. \url{https://doi.org/10.1002/eap.2343}

\leavevmode\vadjust pre{\hypertarget{ref-ferrier2002a}{}}%
Ferrier, S. (2002). Mapping spatial pattern in biodiversity for regional
conservation planning: where to from here? \emph{Systematic Biology},
\emph{51}(2), 331--363. \url{https://doi.org/10.1080/10635150252899806}

\leavevmode\vadjust pre{\hypertarget{ref-firme2023}{}}%
Firme, G. F., Hughes, D. J., Laiolo, L., Roughan, M., Suthers, I. M., \&
Doblin, M. A. (2023). Contrasting phytoplankton composition and primary
productivity in multiple mesoscale eddies along the East Australian
coast. \emph{Deep Sea Research Part I: Oceanographic Research Papers},
\emph{193}, 103952. \url{https://doi.org/10.1016/j.dsr.2022.103952}

\leavevmode\vadjust pre{\hypertarget{ref-foster2013}{}}%
Foster, S. D., Givens, G. H., Dornan, G. J., Dunstan, P. K., \& Darnell,
R. (2013). Modelling biological regions from multi-species and
environmental data. \emph{Environmetrics}, \emph{24}(7), 489--499.
\url{https://doi.org/10.1002/env.2245}

\leavevmode\vadjust pre{\hypertarget{ref-foster2017}{}}%
Foster, S. D., Hill, N. A., \& Lyons, M. (2017). Ecological grouping of
survey sites when sampling artefacts are present. \emph{Journal of the
Royal Statistical Society: Series C (Applied Statistics)}, \emph{66}(5),
1031--1047. \url{https://doi.org/10.1111/rssc.12211}

\leavevmode\vadjust pre{\hypertarget{ref-fromant2022}{}}%
Fromant, A., Eizenberg, Y. H., Poupart, T., Bustamante, P., \& Arnould,
J. P. Y. (2022). Year-round at-sea movements of fairy prions from
southeastern {A}ustralia. \emph{Royal Society Open Science},
\emph{9}(5), 220134. \url{https://doi.org/10.1098/rsos.220134}

\leavevmode\vadjust pre{\hypertarget{ref-ganachaud2014}{}}%
Ganachaud, A., Cravatte, S., Melet, A., Schiller, A., Holbrook, N. J.,
Sloyan, B. M., Widlansky, M. J., Bowen, M., Verron, J., Wiles, P.,
Ridgway, K., Sutton, P., Sprintall, J., Steinberg, C., Brassington, G.,
Cai, W., Davis, R., Gasparin, F., Gourdeau, L., \ldots{} Send, U.
(2014). The Southwest Pacific Ocean circulation and climate experiment
(SPICE). \emph{Journal of Geophysical Research: Oceans}, \emph{119}(11),
7660--7686. \url{https://doi.org/10.1002/2013JC009678}

\leavevmode\vadjust pre{\hypertarget{ref-garcia2022}{}}%
Garcia, V., Schilling, H., Cruz, D., Hawes, S., Everett, J., Roughan,
M., Miskiewicz, A., Pakhomov, E., Jeffs, A., \& Suthers, I. (2022).
Entrainment and development of larval fish assemblages in two
contrasting cold core eddies of the East Australian Current system.
\emph{Marine Ecology Progress Series}, \emph{685}, 1--18.
\url{https://doi.org/10.3354/meps13982}

\leavevmode\vadjust pre{\hypertarget{ref-gotelli2022}{}}%
Gotelli, N. J., Moyes, F., Antão, L. H., Blowes, S. A., Dornelas, M.,
McGill, B. J., Penny, A., Schipper, A. M., Shimadzu, H., Supp, S. R.,
Waldock, C. A., \& Magurran, A. E. (2022). Long-term changes in
temperate marine fish assemblages are driven by a small subset of
species. \emph{Global Change Biology}, \emph{28}(1), 46--53.
\url{https://doi.org/10.1111/gcb.15947}

\leavevmode\vadjust pre{\hypertarget{ref-gruxe9millet2009}{}}%
Grémillet, D., \& Boulinier, T. (2009). Spatial ecology and conservation
of seabirds facing global climate change: a review. \emph{Marine Ecology
Progress Series}, \emph{391}, 121--137.
\url{https://doi.org/10.3354/meps08212}

\leavevmode\vadjust pre{\hypertarget{ref-harris2007}{}}%
Harris, J., Haward, M., Jabour, J., \& Woehler, E. J. (2007). A new
approach to selecting Marine Protected Areas (MPAs) in the Southern
Ocean. \emph{Antarctic Science}, \emph{19}(2), 189--194.
\url{https://doi.org/10.1017/S0954102007000260}

\leavevmode\vadjust pre{\hypertarget{ref-hassler2011}{}}%
Hassler, C. S., Djajadikarta, J. R., Doblin, M. A., Everett, J. D., \&
Thompson, P. A. (2011). Characterisation of water masses and
phytoplankton nutrient limitation in the East Australian Current
separation zone during spring 2008. \emph{Deep Sea Research Part II:
Topical Studies in Oceanography}, \emph{58}(5), 664--677.
\url{https://doi.org/10.1016/j.dsr2.2010.06.008}

\leavevmode\vadjust pre{\hypertarget{ref-haury1978patterns}{}}%
Haury, L., McGowan, J., \& Wiebe, P. (1978). Patterns and processes in
the time-space scales of plankton distributions. In J. Steele (Ed.),
\emph{Spatial pattern in plankton communities} (pp. 277--327). Springer.
\url{https://doi.org/10.1007/978-1-4899-2195-6_12}

\leavevmode\vadjust pre{\hypertarget{ref-hazen2013a}{}}%
Hazen, E. L., Jorgensen, S., Rykaczewski, R. R., Bograd, S. J., Foley,
D. G., Jonsen, I. D., Shaffer, S. A., Dunne, J. P., Costa, D. P.,
Crowder, L. B., \& Block, B. A. (2013). Predicted habitat shifts of
Pacific top predators in a changing climate. \emph{Nature Climate
Change}, \emph{3}(3), 234--238.
\url{https://doi.org/10.1038/nclimate1686}

\leavevmode\vadjust pre{\hypertarget{ref-hazen2013}{}}%
Hazen, E. L., Suryan, R. M., Santora, J. A., Bograd, S., Watanuki, Y.,
\& Wilson, R. P. (2013). Scales and mechanisms of marine hotspot
formation. \emph{Marine Ecology Progress Series}, \emph{487}, 177--183.
\url{https://doi.org/10.3354/meps10477}

\leavevmode\vadjust pre{\hypertarget{ref-hedd2001}{}}%
Hedd, A., Gales, R., \& Brothers, N. (2001). Foraging strategies of shy
albatross {\emph{Thalassarche cauta}} breeding at Albatross Island,
Tasmania, Australia. \emph{Marine Ecology Progress Series}, \emph{224},
267--282. \url{https://doi.org/10.3354/meps224267}

\leavevmode\vadjust pre{\hypertarget{ref-raster}{}}%
Hijmans, R. J. (2022a). \emph{Raster: Geographic data analysis and
modeling}. \url{https://CRAN.R-project.org/package=raster}

\leavevmode\vadjust pre{\hypertarget{ref-terra}{}}%
Hijmans, R. J. (2022b). \emph{Terra: Spatial data analysis}.
\url{https://CRAN.R-project.org/package=terra}

\leavevmode\vadjust pre{\hypertarget{ref-hill2008}{}}%
Hill, K. L., Rintoul, S. R., Coleman, R., \& Ridgway, K. R. (2008). Wind
forced low frequency variability of the East Australia Current.
\emph{Geophysical Research Letters}, \emph{35}(8).
\url{https://doi.org/10.1029/2007GL032912}

\leavevmode\vadjust pre{\hypertarget{ref-hill2017}{}}%
Hill, N. A., Foster, S. D., Duhamel, G., Welsford, D., Koubbi, P., \&
Johnson, C. R. (2017). Model-based mapping of assemblages for ecology
and conservation management: a case study of demersal fish on the
Kerguelen Plateau. \emph{Diversity and Distributions}, \emph{23}(10),
1216--1230. https://doi.org/\url{https://doi.org/10.1111/ddi.12613}

\leavevmode\vadjust pre{\hypertarget{ref-hill2020}{}}%
Hill, N. A., Woolley, S. N. C., Foster, S., Dunstan, P. K., McKinlay,
J., Ovaskainen, O., \& Johnson, C. (2020). Determining marine
bioregions: a comparison of quantitative approaches. \emph{Methods in
Ecology and Evolution}, \emph{11}(10), 1258--1272.
\url{https://doi.org/10.1111/2041-210X.13447}

\leavevmode\vadjust pre{\hypertarget{ref-hobday2010}{}}%
Hobday, A. J. (2010). Ensemble analysis of the future distribution of
large pelagic fishes off Australia. \emph{Progress in Oceanography},
\emph{86}(1), 291--301.
\url{https://doi.org/10.1016/j.pocean.2010.04.023}

\leavevmode\vadjust pre{\hypertarget{ref-hsieh2016}{}}%
Hsieh, T. C., Ma, K. H., \& Chao, A. (2016). iNEXT: an R package for
rarefaction and extrapolation of species diversity (Hill numbers).
\emph{Methods in Ecology and Evolution}, \emph{7}(12), 1451--1456.
https://doi.org/\url{https://doi.org/10.1111/2041-210X.12613}

\leavevmode\vadjust pre{\hypertarget{ref-hunt1987}{}}%
Hunt, G. L., \& Schneider, D. C. (1987). Scale-dependent processes in
the physical and biological environment of marine birds. In J. P.
Croxall (Ed.), \emph{Seabirds: feeding ecology and roles in marine
ecosystem} (pp. 7--41). Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-hyrenbach2007}{}}%
Hyrenbach, K. D., Veit, R. R., Weimerskirch, H., Metzl, N., \& Hunt, G.
L. (2007). Community structure across a large-scale ocean productivity
gradient: marine bird assemblages of the Southern Indian Ocean.
\emph{Deep Sea Research Part I: Oceanographic Research Papers},
\emph{54}(7), 1129--1145.
\url{https://doi.org/10.1016/j.dsr.2007.05.002}

\leavevmode\vadjust pre{\hypertarget{ref-ismar2011}{}}%
Ismar, S. M. H., Phillips, R. A., Rayner, M. J., \& Hauber, M. E.
(2011). Geolocation tracking of the annual migration of adult
Australasian gannets ({\emph{Morus serrator}}) breeding in New Zealand.
\emph{Wilson Journal of Ornithology}, \emph{123}(1), 121--125.
\url{https://doi.org/10.1676/10-072.1}

\leavevmode\vadjust pre{\hypertarget{ref-jaine2014}{}}%
Jaine, F., Rohner, C., Weeks, S., Couturier, L., Bennett, M., Townsend,
K., \& Richardson, A. (2014). Movements and habitat use of reef manta
rays off eastern Australia: offshore excursions, deep diving and eddy
affinity revealed by satellite telemetry. \emph{Marine Ecology Progress
Series}, \emph{510}, 73--86. \url{https://doi.org/10.3354/meps10910}

\leavevmode\vadjust pre{\hypertarget{ref-johnson2021}{}}%
Johnson, B., \& Colombelli-Négrel, D. (2021). Breeding success in
Southern Australian little penguins is negatively correlated with high
wind speeds and sea surface temperatures. \emph{Ornithological
Applications}, \emph{123}(1), duaa062.
\url{https://doi.org/10.1093/ornithapp/duaa062}

\leavevmode\vadjust pre{\hypertarget{ref-johnson2011}{}}%
Johnson, C. R., Banks, S. C., Barrett, N. S., Cazassus, F., Dunstan, P.
K., Edgar, G. J., Frusher, S. D., Gardner, C., Haddon, M., Helidoniotis,
F., Hill, K. L., Holbrook, N. J., Hosie, G. W., Last, P. R., Ling, S.
D., Melbourne-Thomas, J., Miller, K., Pecl, G. T., Richardson, A. J.,
\ldots{} Taw, N. (2011). Climate change cascades: Shifts in
oceanography, species' ranges and subtidal marine community dynamics in
eastern Tasmania. \emph{Journal of Experimental Marine Biology and
Ecology}, \emph{400}(1), 17--32.
\url{https://doi.org/10.1016/j.jembe.2011.02.032}

\leavevmode\vadjust pre{\hypertarget{ref-karpouzi2007}{}}%
Karpouzi, V., Watson, R., \& Pauly, D. (2007). Modelling and mapping
resource overlap between seabirds and fisheries on a global scale: a
preliminary assessment. \emph{Marine Ecology Progress Series},
\emph{343}, 87--99. \url{https://doi.org/10.3354/meps06860}

\leavevmode\vadjust pre{\hypertarget{ref-kessler2013}{}}%
Kessler, W. S., \& Cravatte, S. (2013). Mean circulation of the Coral
Sea. \emph{Journal of Geophysical Research: Oceans}, \emph{118}(12),
6385--6410. \url{https://doi.org/10.1002/2013JC009117}

\leavevmode\vadjust pre{\hypertarget{ref-kreft2010}{}}%
Kreft, H., \& Jetz, W. (2010). A framework for delineating
biogeographical regions based on species distributions. \emph{Journal of
Biogeography}, \emph{37}(11), 2029--2053.
\url{https://doi.org/10.1111/j.1365-2699.2010.02375.x}

\leavevmode\vadjust pre{\hypertarget{ref-lavers2014}{}}%
Lavers, J. L., Miller, M. G. R., Carter, M. J., Swann, G., \& Clarke, R.
H. (2014). Predicting the spatial distribution of a seabird community to
identify priority conservation areas in the Timor Sea.
\emph{Conservation Biology}, \emph{28}(6), 1699--1709.
\url{https://doi.org/10.1111/cobi.12324}

\leavevmode\vadjust pre{\hypertarget{ref-lee2019good}{}}%
Lee, M. R., Powell, J. R., Oberle, B., Cornwell, W. K., Lyons, M., Rigg,
J. L., \& Zanne, A. E. (2019). Good neighbors aplenty: Fungal endophytes
rarely exhibit competitive exclusion patterns across a span of woody
habitats. \emph{Ecology}, \emph{100}(9), e02790.
\url{https://doi.org/10.1002/ecy.2790}

\leavevmode\vadjust pre{\hypertarget{ref-lenoir2020}{}}%
Lenoir, J., Bertrand, R., Comte, L., Bourgeaud, L., Hattab, T.,
Murienne, J., \& Grenouillet, G. (2020). Species better track climate
warming in the oceans than on land. \emph{Nature Ecology \& Evolution},
\emph{4}(8), 1044--1059. \url{https://doi.org/10.1038/s41559-020-1198-2}

\leavevmode\vadjust pre{\hypertarget{ref-li2022}{}}%
Li, J., Roughan, M., \& Kerry, C. (2022). Drivers of ocean warming in
the western boundary currents of the Southern Hemisphere. \emph{Nature
Climate Change}, \emph{12}(10), 901--909.
\url{https://doi.org/10.1038/s41558-022-01473-8}

\leavevmode\vadjust pre{\hypertarget{ref-longhurst2007}{}}%
Longhurst, A. R. (2007). \emph{Ecological geography of the sea} (2nd
ed). Academic Press.

\leavevmode\vadjust pre{\hypertarget{ref-lyons2017}{}}%
Lyons, M. B., Foster, S. D., \& Keith, D. A. (2017). Simultaneous
vegetation classification and mapping at large spatial scales.
\emph{Journal of Biogeography}, \emph{44}(12), 2891--2902.
\url{https://doi.org/10.1111/jbi.13088}

\leavevmode\vadjust pre{\hypertarget{ref-marchant1990}{}}%
Marchant, R. (1990). Robustness of classification and ordination
techniques applied to macroinvertebrate communities from the La Trobe
River, Victoria. \emph{Marine and Freshwater Research}, \emph{41}(4),
493--504. \url{https://doi.org/10.1071/MF9900493}

\leavevmode\vadjust pre{\hypertarget{ref-marchant1990handbook}{}}%
Marchant, S., \& Higgins, P. J. (1990). \emph{Handbook of australian,
new zealand \& antarctic birds. Vol. 1, ratites to ducks, p. AB}. Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mason2018}{}}%
Mason, C., Alderman, R., McGowan, J., Possingham, H. P., Hobday, A. J.,
Sumner, M., \& Shaw, J. (2018). Telemetry reveals existing marine
protected areas are worse than random for protecting the foraging
habitat of threatened shy albatross ({\emph{Thalassarche cauta}}).
\emph{Diversity and Distributions}, \emph{24}(12), 1744--1755.
\url{https://doi.org/10.1111/ddi.12830}

\leavevmode\vadjust pre{\hypertarget{ref-mata2006}{}}%
Mata, M. M., Wijffels, S. E., Church, J. A., \& Tomczak, M. (2006). Eddy
shedding and energy conversions in the East Australian Current.
\emph{Journal of Geophysical Research: Oceans}, \emph{111}, C09034.
\url{https://doi.org/10.1029/2006JC003592}

\leavevmode\vadjust pre{\hypertarget{ref-mcduie2015}{}}%
McDuie, F., Weeks, S. J., Miller, M. G. R., \& Congdon, B. C. (2015).
Breeding tropical shearwaters use distant foraging sites when
self-provisioning. \emph{Marine Ornithology}, \emph{43}, 123--129.

\leavevmode\vadjust pre{\hypertarget{ref-melo2005}{}}%
Melo, A. S. (2005). Effects of taxonomic and numeric resolution on the
ability to detect ecological patterns at a local scale using stream
macroinvertebrates. \emph{Archiv Für Hydrobiologie}, \emph{164}(3),
309--323. \url{https://doi.org/10.1127/0003-9136/2005/0164-0309}

\leavevmode\vadjust pre{\hypertarget{ref-miller2018}{}}%
Miller, M. G. R., Silva, F. R. O., Machovsky-Capuska, G. E., \& Congdon,
B. C. (2018). Sexual segregation in tropical seabirds: drivers of
sex-specific foraging in the brown booby {\emph{Sula leucogaster}}.
\emph{Journal of Ornithology}, \emph{159}(2), 425--437.
\url{https://doi.org/10.1007/s10336-017-1512-1}

\leavevmode\vadjust pre{\hypertarget{ref-morrone2018}{}}%
Morrone, J. J. (2018). The spectre of biogeographical regionalization.
\emph{Journal of Biogeography}, \emph{45}(2), 282--288.
\url{https://doi.org/10.1111/jbi.13135}

\leavevmode\vadjust pre{\hypertarget{ref-morrone2023biogeographical}{}}%
Morrone, J. J. (2023). Why biogeographical transition zones matter.
\emph{Journal of Biogeography}. \url{https://doi.org/10.1111/jbi.14632}

\leavevmode\vadjust pre{\hypertarget{ref-mott2018}{}}%
Mott, R., \& Clarke, R. H. (2018). Systematic review of geographic
biases in the collection of at-sea distribution data for seabirds.
\emph{Emu - Austral Ornithology}, \emph{118}(3), 235--246.
\url{https://doi.org/10.1080/01584197.2017.1416957}

\leavevmode\vadjust pre{\hypertarget{ref-niella2022}{}}%
Niella, Y., Butcher, P., Holmes, B., Barnett, A., \& Harcourt, R.
(2022). Forecasting intraspecific changes in distribution of a
wide-ranging marine predator under climate change. \emph{Oecologia},
\emph{198}(1), 111--124.
\url{https://doi.org/10.1007/s00442-021-05075-7}

\leavevmode\vadjust pre{\hypertarget{ref-niella2020}{}}%
Niella, Y., Smoothey, A., Peddemors, V., \& Harcourt, R. (2020).
Predicting changes in distribution of a large coastal shark in the face
of the strengthening East Australian Current. \emph{Marine Ecology
Progress Series}, \emph{642}, 163--177.
\url{https://doi.org/10.3354/meps13322}

\leavevmode\vadjust pre{\hypertarget{ref-norman1992distribution}{}}%
Norman, F. (1992). Distribution and abundance of seabirds off {P}hillip
{I}sland and within {P}ort {P}hillip {B}ay, {V}ictoria, 1986--1988.
\emph{Emu - Austral Ornithology}, \emph{91}(5), 377--394.

\leavevmode\vadjust pre{\hypertarget{ref-ohara2020}{}}%
O'Hara, T. D., Williams, A., Althaus, F., Ross, A. S., \& Bax, N. J.
(2020). Regional-scale patterns of deep seafloor biodiversity for
conservation assessment. \emph{Diversity and Distributions},
\emph{26}(4), 479--494. \url{https://doi.org/10.1111/ddi.13034}

\leavevmode\vadjust pre{\hypertarget{ref-ohara2020a}{}}%
O'Hara, T. D., Williams, A., Woolley, S. N. C., Nau, A. W., \& Bax, N.
J. (2020). Deep-sea temperate-tropical faunal transition across uniform
environmental gradients. \emph{Deep Sea Research Part I: Oceanographic
Research Papers}, \emph{161}, 103283.
\url{https://doi.org/10.1016/j.dsr.2020.103283}

\leavevmode\vadjust pre{\hypertarget{ref-oke2000}{}}%
Oke, P. R., \& Middleton, J. H. (2000). Topographically induced
upwelling off eastern Australia. \emph{Journal of Physical
Oceanography}, \emph{30}(3), 512--531.
\url{https://doi.org/10.1175/1520-0485(2000)030\%3C0512:TIUOEA\%3E2.0.CO;2}

\leavevmode\vadjust pre{\hypertarget{ref-oke2019a}{}}%
Oke, P. R., Pilo, G. S., Ridgway, K., Kiss, A., \& Rykova, T. (2019). A
search for the Tasman Front. \emph{Journal of Marine Systems},
\emph{199}, 103217. \url{https://doi.org/10.1016/j.jmarsys.2019.103217}

\leavevmode\vadjust pre{\hypertarget{ref-oke2019}{}}%
Oke, P. R., Roughan, M., Cetina-Heredia, P., Pilo, G. S., Ridgway, K.
R., Rykova, T., Archer, M. R., Coleman, R. C., Kerry, C. G., Rocha, C.,
Schaeffer, A., \& Vitarelli, E. (2019). Revisiting the circulation of
the East Australian Current: its path, separation, and eddy field.
\emph{Progress in Oceanography}, \emph{176}, 102139.
\url{https://doi.org/10.1016/j.pocean.2019.102139}

\leavevmode\vadjust pre{\hypertarget{ref-oliver2017}{}}%
Oliver, E. C. J., Benthuysen, J. A., Bindoff, N. L., Hobday, A. J.,
Holbrook, N. J., Mundy, C. N., \& Perkins-Kirkpatrick, S. E. (2017). The
unprecedented 2015/16 Tasman Sea marine heatwave. \emph{Nature
Communications}, \emph{8}(1), 16101.
\url{https://doi.org/10.1038/ncomms16101}

\leavevmode\vadjust pre{\hypertarget{ref-oliver2018}{}}%
Oliver, E. C. J., Lago, V., Hobday, A. J., Holbrook, N. J., Ling, S. D.,
\& Mundy, C. N. (2018). Marine heatwaves off eastern Tasmania: trends,
interannual variability, and predictability. \emph{Progress in
Oceanography}, \emph{161}, 116--130.
\url{https://doi.org/10.1016/j.pocean.2018.02.007}

\leavevmode\vadjust pre{\hypertarget{ref-pantiukhin2021}{}}%
Pantiukhin, D., Piepenburg, D., Hansen, M. L. S., \& Kraan, C. (2021).
Data-driven bioregionalization: a seascape-scale study of macrobenthic
communities in the Eurasian Arctic. \emph{Journal of Biogeography},
\emph{48}(11), 2877--2890. \url{https://doi.org/10.1111/jbi.14247}

\leavevmode\vadjust pre{\hypertarget{ref-sf}{}}%
Pebesma, E. (2018). {Simple features for R: standardized support for
spatial vector data}. \emph{{The R Journal}}, \emph{10}(1), 439--446.
\url{https://doi.org/10.32614/RJ-2018-009}

\leavevmode\vadjust pre{\hypertarget{ref-peck2004}{}}%
Peck, D., Smithers, B., Krockenberger, A., \& Congdon, B. (2004). Sea
surface temperature constrains wedge-tailed shearwater foraging success
within breeding seasons. \emph{Marine Ecology Progress Series},
\emph{281}, 259--266. \url{https://doi.org/10.3354/meps281259}

\leavevmode\vadjust pre{\hypertarget{ref-puxe9ron2010}{}}%
Péron, C., Authier, M., Barbraud, C., Delord, K., Besson, D., \&
Weimerskirch, H. (2010). Interdecadal changes in at-sea distribution and
abundance of subantarctic seabirds along a latitudinal gradient in the
Southern Indian Ocean. \emph{Global Change Biology}, \emph{16}(7),
1895--1909. \url{https://doi.org/10.1111/j.1365-2486.2010.02169.x}

\leavevmode\vadjust pre{\hypertarget{ref-phillips2022}{}}%
Phillips, L. R., Malan, N., Roughan, M., Harcourt, R., Jonsen, I., Cox,
M., Brierley, A. S., Slip, D., Wilkins, A., \& Carroll, G. (2022).
Coastal seascape variability in the intensifying {E}ast {A}ustralian
{C}urrent southern extension. \emph{Frontiers in Marine Science},
\emph{9}, 925123.
\url{https://www.frontiersin.org/articles/10.3389/fmars.2022.925123}

\leavevmode\vadjust pre{\hypertarget{ref-pinsky2019}{}}%
Pinsky, M. L., Eikeset, A. M., McCauley, D. J., Payne, J. L., \& Sunday,
J. M. (2019). Greater vulnerability to warming of marine versus
terrestrial ectotherms. \emph{Nature}, \emph{569}(7754), 108--111.
\url{https://doi.org/10.1038/s41586-019-1132-4}

\leavevmode\vadjust pre{\hypertarget{ref-pocklington1979}{}}%
Pocklington, R. (1979). An oceanographic interpretation of seabird
distributions in the Indian Ocean. \emph{Marine Biology}, \emph{51}(1),
9--21. \url{https://doi.org/10.1007/BF00389026}

\leavevmode\vadjust pre{\hypertarget{ref-poloczanska2007}{}}%
Poloczanska, E. S., Babcock, R. C., Butler, A., Hobday, A. J.,
Hoegh-Guldberg, O., Kunz, T. J., Matear, R., Milton, D., Okey, T. A., \&
Richardson, A. J. (2007). Climate change and Australian marine life. In
R. N. Gibson, R. J. A. Atkinson, \& J. D. M. Gordon (Eds.),
\emph{Oceanography and marine biology: an annual review} (Vol. 45, pp.
407--478). CRC Press, Taylor \& Francis Group.

\leavevmode\vadjust pre{\hypertarget{ref-poloczanska2013}{}}%
Poloczanska, E. S., Brown, C. J., Sydeman, W. J., Kiessling, W.,
Schoeman, D. S., Moore, P. J., Brander, K., Bruno, J. F., Buckley, L.
B., Burrows, M. T., Duarte, C. M., Halpern, B. S., Holding, J., Kappel,
C. V., O'Connor, M. I., Pandolfi, J. M., Parmesan, C., Schwing, F.,
Thompson, S. A., \& Richardson, A. J. (2013). Global imprint of climate
change on marine life. \emph{Nature Climate Change}, \emph{3}(10),
919--925. \url{https://doi.org/10.1038/nclimate1958}

\leavevmode\vadjust pre{\hypertarget{ref-ponti2023}{}}%
Ponti, R., \& Sannolo, M. (2023). The importance of including phenology
when modelling species ecological niche. \emph{Ecography},
\emph{2023}(4), e06143. \url{https://doi.org/10.1111/ecog.06143}

\leavevmode\vadjust pre{\hypertarget{ref-priddel2014}{}}%
Priddel, D., Carlile, N., Portelli, D., Kim, Y., O'Neill, L.,
Bretagnolle, V., Ballance, L. T., Phillips, R. A., Pitman, R. L., \&
Raynei, M. J. (2014). Pelagic distribution of Gould's petrel
({\emph{Pterodroma leucoptera}}): linking shipboard and onshore
observations with remote-tracking data. \emph{Emu - Austral
Ornithology}, \emph{114}(4), 360--370.
\url{https://doi.org/10.1071/MU14021}

\leavevmode\vadjust pre{\hypertarget{ref-quillfeldt2015}{}}%
Quillfeldt, P., Cherel, Y., Delord, K., \& Weimerkirch, H. (2015). Cool,
cold or colder? Spatial segregation of prions and blue petrels is
explained by differences in preferred sea surface temperatures.
\emph{Biology Letters}, \emph{11}(4), 20141090.
\url{https://doi.org/10.1098/rsbl.2014.1090}

\leavevmode\vadjust pre{\hypertarget{ref-r2022}{}}%
R Core Team. (2022). \emph{R: A language and environment for statistical
computing}. R Foundation for Statistical Computing.
\url{https://www.R-project.org/}

\leavevmode\vadjust pre{\hypertarget{ref-ramuxedrez2017}{}}%
Ramírez, F., Afán, I., Davis, L. S., \& Chiaradia, A. (2017). Climate
impacts on global hot spots of marine biodiversity. \emph{Science
Advances}, \emph{3}(2), e1601198.
\url{https://doi.org/10.1126/sciadv.1601198}

\leavevmode\vadjust pre{\hypertarget{ref-raymond2010}{}}%
Raymond, B., Shaffer, S. A., Sokolov, S., Woehler, E. J., Costa, D. P.,
Einoder, L., Hindell, M., Hosie, G., Pinkerton, M., Sagar, P. M., Scott,
D., Smith, A., Thompson, D. R., Vertigan, C., \& Weimerskirch, H.
(2010). Shearwater foraging in the Southern Ocean: the roles of prey
availability and winds. \emph{PLoS ONE}, \emph{5}(6), e10960.
\url{https://doi.org/10.1371/journal.pone.0010960}

\leavevmode\vadjust pre{\hypertarget{ref-receveur2020}{}}%
Receveur, A., Vourey, E., Lebourges-Dhaussy, A., Menkes, C., Ménard, F.,
\& Allain, V. (2020). Biogeography of micronekton assemblages in the
Natural Park of the Coral Sea. \emph{Frontiers in Marine Science},
\emph{7}, 449. \url{https://doi.org/10.3389/fmars.2020.00449}

\leavevmode\vadjust pre{\hypertarget{ref-reid2023}{}}%
Reid, K., Baker, G. B., \& Woehler, E. J. (2023). An ecological risk
assessment for the impacts of offshore wind farms on birds in Australia.
\emph{Austral Ecology}, \emph{48}(2), 418--439.
\url{https://doi.org/10.1111/aec.13278}

\leavevmode\vadjust pre{\hypertarget{ref-reid2000}{}}%
Reid, T. A., \& Hindell, M. A. (2000). Coarse-scale relationships
between seabirds and zooplankton off south-eastern Tasmania.
\emph{Marine and Freshwater Research}, \emph{51}(8), 789--798.
\url{https://doi.org/10.1071/MF98160}

\leavevmode\vadjust pre{\hypertarget{ref-reisinger2022}{}}%
Reisinger, R. R., Corney, S., Raymond, B., Lombard, A. T., Bester, M.
N., Crawford, R. J. M., Davies, D., de Bruyn, P. J. N., Dilley, B. J.,
Kirkman, S. P., Makhado, A. B., Ryan, P. G., Schoombie, S., Stevens, K.
L., Tosh, C. A., Wege, M., Whitehead, T. O., Sumner, M. D., Wotherspoon,
S., \ldots{} Pistorius, P. A. (2022). Habitat model forecasts suggest
potential redistribution of marine predators in the southern Indian
Ocean. \emph{Diversity and Distributions}, \emph{28}(1), 142--159.
\url{https://doi.org/10.1111/ddi.13447}

\leavevmode\vadjust pre{\hypertarget{ref-reygondeau2013}{}}%
Reygondeau, G., Longhurst, A., Martinez, E., Beaugrand, G., Antoine, D.,
\& Maury, O. (2013). Dynamic biogeochemical provinces in the global
ocean. \emph{Global Biogeochemical Cycles}, \emph{27}(4), 1046--1058.
\url{https://doi.org/10.1002/gbc.20089}

\leavevmode\vadjust pre{\hypertarget{ref-reygondeau2012}{}}%
Reygondeau, G., Maury, O., Beaugrand, G., Fromentin, J. M., Fonteneau,
A., \& Cury, P. (2012). Biogeography of tuna and billfish communities.
\emph{Journal of Biogeography}, \emph{39}(1), 114--129.
\url{https://doi.org/10.1111/j.1365-2699.2011.02582.x}

\leavevmode\vadjust pre{\hypertarget{ref-ribic1997}{}}%
Ribic, C., Ainley, D., \& Spear, L. (1997). Scale-related
seabird-environmental relationships in Pacific equatorial waters, with
reference to El Niño-Southern Oscillation events. \emph{Marine Ecology
Progress Series}, \emph{156}, 183--203.
\url{https://doi.org/10.3354/meps156183}

\leavevmode\vadjust pre{\hypertarget{ref-ridgway2003a}{}}%
Ridgway, K. R., \& Dunn, J. R. (2003). Mesoscale structure of the mean
East Australian Current System and its relationship with topography.
\emph{Progress in Oceanography}, \emph{56}(2), 189--222.
\url{https://doi.org/10.1016/S0079-6611(03)00004-1}

\leavevmode\vadjust pre{\hypertarget{ref-santora2013}{}}%
Santora, J., \& Veit, R. (2013). Spatio-temporal persistence of top
predator hotspots near the Antarctic Peninsula. \emph{Marine Ecology
Progress Series}, \emph{487}, 287--304.
\url{https://doi.org/10.3354/meps10350}

\leavevmode\vadjust pre{\hypertarget{ref-scales2017}{}}%
Scales, K. L., Hazen, E. L., Jacox, M. G., Edwards, C. A., Boustany, A.
M., Oliver, M. J., \& Bograd, S. J. (2017). Scale of inference: on the
sensitivity of habitat models for wide-ranging marine predators to the
resolution of environmental data. \emph{Ecography}, \emph{40}(1),
210--220. \url{https://doi.org/10.1111/ecog.02272}

\leavevmode\vadjust pre{\hypertarget{ref-scales2014a}{}}%
Scales, K. L., Miller, P. I., Embling, C. B., Ingram, S. N., Pirotta,
E., \& Votier, S. C. (2014). Mesoscale fronts as foraging habitats:
composite front mapping reveals oceanographic drivers of habitat use for
a pelagic seabird. \emph{Journal of The Royal Society Interface},
\emph{11}(100), 20140679. \url{https://doi.org/10.1098/rsif.2014.0679}

\leavevmode\vadjust pre{\hypertarget{ref-scales2014}{}}%
Scales, K. L., Miller, P. I., Hawkes, L. A., Ingram, S. N., Sims, D. W.,
\& Votier, S. C. (2014). On the front line: frontal zones as priority
at-sea conservation areas for mobile marine vertebrates. \emph{Journal
of Applied Ecology}, \emph{51}(6), 1575--1583.
\url{https://doi.org/10.1111/1365-2664.12330}

\leavevmode\vadjust pre{\hypertarget{ref-biology2002}{}}%
Schreiber, E. A., \& Burger, J. (Eds.). (2002). \emph{Biology of marine
birds}. CRC Press.

\leavevmode\vadjust pre{\hypertarget{ref-sherman1994}{}}%
Sherman, K. (1994). Sustainability, biomass yields, and health of
coastal ecosystems: an ecological perspective. \emph{Marine Ecology
Progress Series}, \emph{112}, 277--301.
\url{https://doi.org/10.3354/meps112277}

\leavevmode\vadjust pre{\hypertarget{ref-sojitra2022}{}}%
Sojitra, M., Woehler, E. J., Lea, M.-A., \& Wotherspoon, S. (2022).
Multi-decadal changes in the at-sea distribution and abundance of
black-browed and light-mantled sooty albatrosses in the {S}outhwest
{P}acific {O}cean. \emph{ICES Journal of Marine Science}, fsac197.
\url{https://doi.org/10.1093/icesjms/fsac197}

\leavevmode\vadjust pre{\hypertarget{ref-souza2023}{}}%
Souza, J. S., \& Santos, L. N. (2023). Resident species, not immigrants,
drive reorganization of estuarine fish assemblages in response to
warming. \emph{Ecology}, \emph{104}(5), e3987.
\url{https://doi.org/10.1002/ecy.3987}

\leavevmode\vadjust pre{\hypertarget{ref-spalding2007}{}}%
Spalding, M. D., Fox, H. E., Allen, G. R., Davidson, N., Ferdaña, Z. A.,
Finlayson, M., Halpern, B. S., Jorge, M. A., Lombana, A., Lourie, S. A.,
Martin, K. D., McManus, E., Molnar, J., Recchia, C. A., \& Robertson, J.
(2007). Marine {E}coregions of the {W}orld: A bioregionalization of
coastal and shelf areas. \emph{BioScience}, \emph{57}(7), 573--583.
\url{https://doi.org/10.1641/B570707}

\leavevmode\vadjust pre{\hypertarget{ref-sun2022}{}}%
Sun, C., Zhang, A., Jin, B., Wang, X., Zhang, X., \& Zhang, L. (2022).
Seasonal variability of eddy kinetic energy in the north indian ocean.
\emph{Frontiers in Marine Science}, \emph{9}, 1032699.
\url{https://doi.org/10.3389/fmars.2022.1032699}

\leavevmode\vadjust pre{\hypertarget{ref-suthers2011}{}}%
Suthers, I. M., Young, J. W., Baird, M. E., Roughan, M., Everett, J. D.,
Brassington, G. B., Byrne, M., Condie, S. A., Hartog, J. R., Hassler, C.
S., Hobday, A. J., Holbrook, N. J., Malcolm, H. A., Oke, P. R.,
Thompson, P. A., \& Ridgway, K. (2011). The strengthening East
Australian Current, its eddies and biological effects {\textemdash} an
introduction and overview. \emph{Deep Sea Research Part II: Topical
Studies in Oceanography}, \emph{58}(5), 538--546.
\url{https://doi.org/10.1016/j.dsr2.2010.09.029}

\leavevmode\vadjust pre{\hypertarget{ref-sutton2017a}{}}%
Sutton, T. T., Clark, M. R., Dunn, D. C., Halpin, P. N., Rogers, A. D.,
Guinotte, J., Bograd, S. J., Angel, M. V., Perez, J. A. A., Wishner, K.,
Haedrich, R. L., Lindsay, D. J., Drazen, J. C., Vereshchaka, A.,
Piatkowski, U., Morato, T., Błachowiak-Samołyk, K., Robison, B. H.,
Gjerde, K. M., \ldots{} Heino, M. (2017). A global biogeographic
classification of the mesopelagic zone. \emph{Deep Sea Research Part I:
Oceanographic Research Papers}, \emph{126}, 85--102.
\url{https://doi.org/10.1016/j.dsr.2017.05.006}

\leavevmode\vadjust pre{\hypertarget{ref-sydeman2012}{}}%
Sydeman, W., Thompson, S., \& Kitaysky, A. (2012). Seabirds and climate
change: roadmap for the future. \emph{Marine Ecology Progress Series},
\emph{454}, 107--117. \url{https://doi.org/10.3354/meps09806}

\leavevmode\vadjust pre{\hypertarget{ref-taylor2021}{}}%
Taylor, G., Elliott, G., Walker, K., \& Bose, S. (2021). Year-round
distribution, breeding cycle, and activity of white-headed petrels
({\emph{Pterodroma lessonii}}) nesting on adams island, auckland
islands. \emph{Notornis}, \emph{67}, 369--386.

\leavevmode\vadjust pre{\hypertarget{ref-tremblay2009}{}}%
Tremblay, Y., Bertrand, S., Henry, R., Kappes, M., Costa, D., \&
Shaffer, S. (2009). Analytical approaches to investigating
seabird--environment interactions: a review. \emph{Marine Ecology
Progress Series}, \emph{391}, 153--163.
\url{https://doi.org/10.3354/meps08146}

\leavevmode\vadjust pre{\hypertarget{ref-vilhena2015}{}}%
Vilhena, D. A., \& Antonelli, A. (2015). A network approach for
identifying and delimiting biogeographical regions. \emph{Nature
Communications}, \emph{6}(1), 6848.
\url{https://doi.org/10.1038/ncomms7848}

\leavevmode\vadjust pre{\hypertarget{ref-warton2015}{}}%
Warton, D. I., Foster, S. D., De'ath, G., Stoklosa, J., \& Dunstan, P.
K. (2015). Model-based thinking for community ecology. \emph{Plant
Ecology}, \emph{216}(5), 669--682.
\url{https://doi.org/10.1007/s11258-014-0366-3}

\leavevmode\vadjust pre{\hypertarget{ref-waters2008marine}{}}%
Waters, J. M. (2008). Marine biogeographical disjunction in temperate
{A}ustralia: Historical landbridge, contemporary currents, or both?
\emph{Diversity and Distributions}, \emph{14}(4), 692--700.
\url{https://doi.org/10.1111/j.1472-4642.2008.00481.x}

\leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2007}{}}%
Weimerskirch, H. (2007). Are seabirds foraging for unpredictable
resources? \emph{Deep Sea Research Part II: Topical Studies in
Oceanography}, \emph{54}(3-4), 211--223.
\url{https://doi.org/10.1016/j.dsr2.2006.11.013}

\leavevmode\vadjust pre{\hypertarget{ref-weimerskirch2020}{}}%
Weimerskirch, H., De Grissac, S., Ravache, A., Prudor, A., Corbeau, A.,
Congdon, B., McDuie, F., Bourgeois, K., Dromzée, S., Butscher, J.,
Menkes, C., Allain, V., Vidal, E., Jaeger, A., \& Borsa, P. (2020).
At-sea movements of wedge-tailed shearwaters during and outside the
breeding season from four colonies in New Caledonia. \emph{Marine
Ecology Progress Series}, \emph{633}, 225--238.
\url{https://doi.org/10.3354/meps13171}

\leavevmode\vadjust pre{\hypertarget{ref-whittaker2005}{}}%
Whittaker, R. J., Araújo, M. B., Jepson, P., Ladle, R. J., Watson, J. E.
M., \& Willis, K. J. (2005). Conservation Biogeography: assessment and
prospect. \emph{Diversity and Distributions}, \emph{11}(1), 3--23.
\url{https://doi.org/10.1111/j.1366-9516.2005.00143.x}

\leavevmode\vadjust pre{\hypertarget{ref-woehler2023}{}}%
Woehler, E., \& Hobday, A. (2023). Impacts of marine heatwaves may be
mediated by seabird life history strategies. \emph{Marine Ecology
Progress Series}, \emph{HEAT}. \url{https://doi.org/10.3354/meps14333}

\leavevmode\vadjust pre{\hypertarget{ref-woehler2003}{}}%
Woehler, E., Raymond, B., \& Watts, D. (2003). Decadal-scale seabird
assemblages in Prydz Bay, East Antarctica. \emph{Marine Ecology Progress
Series}, \emph{251}, 299--310. \url{https://doi.org/10.3354/meps251299}

\leavevmode\vadjust pre{\hypertarget{ref-woolley2020}{}}%
Woolley, S. N. C., Foster, S. D., Bax, N. J., Currie, J. C., Dunn, D.
C., Hansen, C., Hill, N. A., O'Hara, T. D., Ovaskainen, O., Sayre, R.,
Vanhatalo, J. P., \& Dunstan, P. K. (2020). Bioregions in marine
environments: Combining biological and environmental data for management
and scientific understanding. \emph{BioScience}, \emph{70}(1), 48--59.
\url{https://doi.org/10.1093/biosci/biz133}

\leavevmode\vadjust pre{\hypertarget{ref-ecomix}{}}%
Woolley, S., Foster, S., \& Dunstan, P. (2022). \emph{Ecomix: Fitting
finite mixture models to ecological data}.
\url{https://github.com/skiptoniam/ecomix}

\leavevmode\vadjust pre{\hypertarget{ref-yang2016}{}}%
Yang, H., Lohmann, G., Wei, W., Dima, M., Ionita, M., \& Liu, J. (2016).
Intensification and poleward shift of subtropical western boundary
currents in a warming climate. \emph{Journal of Geophysical Research:
Oceans}, \emph{121}(7), 4928--4945.
\url{https://doi.org/10.1002/2015JC011513}

\leavevmode\vadjust pre{\hypertarget{ref-young2011}{}}%
Young, J. W., Hobday, A. J., Campbell, R. A., Kloser, R. J., Bonham, P.
I., Clementson, L. A., \& Lansdell, M. J. (2011). The biological
oceanography of the East Australian Current and surrounding waters in
relation to tuna and billfish catches off eastern Australia. \emph{Deep
Sea Research Part II: Topical Studies in Oceanography}, \emph{58}(5),
720--733. \url{https://doi.org/10.1016/j.dsr2.2010.10.005}

\end{CSLReferences}

\newpage

\hypertarget{tables-and-figures}{%
\section*{Tables and Figures}\label{tables-and-figures}}
\addcontentsline{toc}{section}{Tables and Figures}

\begin{itemize}
\tightlist
\item
  \textbf{TABLE 1.} Environmental data used in the Region of Common
  Profile models applied to seabird data off eastern Australia, their
  spatial/temporal resolutions, units, and sources
\item
  \textbf{TABLE 2.} Full (a) and best (b) Region of Common Profile
  Models applied to seabird data off eastern Australia, for both data
  types used (presence-absence, based on a Bernoulli model; abundance,
  based on a Negative Binomial model). Refer to Table 1 for the
  environmental data acronyms
\item
  \textbf{FIGURE 1.} Study area showing the main ocean currents (a), and
  seabird records made by season on top of the 1° latitude × 1°
  longitude grid cells (b). In (a), the East Australian Current (EAC)
  system is highlighted with its acronyms in bold and drawn in blue. The
  South Equatorial Current (SEC) and its branches, the North Caledonian
  Jet (NCJ) and South Caledonian Jet (SCJ), are drawn in red. The EAC
  `eastern extension' is an eddy field (blue shaded area) often called
  the `Tasman Front', from which the East Auckland Current originates
  (EAUC, in green). Figure (a) is adapted after Oke, Roughan, et
  al.~(2019); for colour references see the online version
\item
  \textbf{FIGURE 2.} Partial plots for the retained covariates in the
  spring (a) presence-absence and (b) abundance best models. The plot
  shows the fitted probability of belonging to a Region of Common
  Profiles (RCP) against the environmental value. All other predictors
  are held at their mean values. Results for other seasons and models
  can be found in the Supplementary Material (Fig. S5). Refer to Table 1
  for the environmental data acronyms
\item
  \textbf{FIGURE 3.} Spatialized point predictions (hard-classes) for
  seabird assemblages (Region of Common Profile; RCP), by season, based
  on presence-absence (a) and abundance (b) models, off eastern
  Australia. Each grid was assigned to the RCP group with highest
  membership probability
\item
  \textbf{FIGURE 4.} Predicted probability membership of each seabird
  assemblage (Region of Common Profiles; RCP) and grid, off eastern
  Australia, from spring presence-absence (a) and abundance (b) models.
  The central column, `mean', corresponds to the point prediction and
  Bayesian boot-strapped, lower and upper confidence intervals (CI), on
  its sides. Note that, in case of two RCPs, RCP2 probablility equals to
  (1 - RCP1 probability). Results for other seasons and models can be
  found in the Supplementary Material (Fig. S6)
\item
  \textbf{FIGURE 5.} Species profiles of seabird assemblages (Region of
  Common Profile; RCP) off eastern Australia in spring. Values are the
  average and 95\% confidence intervals of probability of occurrence (a)
  and predicted mean abundance (b) for each species, based on 1000
  Bayesian bootstraps. Values in (b) were log10-transformed to
  accommodate the high variation between species. Results for other
  seasons and models can be found in the Supplementary Material (Fig.
  S7)
\end{itemize}

\begin{landscape}

\begin{table}

\caption{\label{tab:table1} }
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}llll>{\raggedright\arraybackslash}p{3cm}}
\toprule
\textbf{Variable} & \textbf{Abbreviation} & \textbf{Units} & \textbf{Spatial.resolution} & \textbf{Temporal.resolution} & \textbf{Source}\\
\midrule
Sea Surface Temperature & sst & °Celsius & 0.25° × 0.25° & daily & Huang et al. (2021)${^a}$\\
Sea Surface Temperature gradient & sst\_grad & Δ°Celsius & 0.25° × 0.25° & daily & Derived from Huang et al. (2021)${^a}$\\
Sea Surface Salinity & sss & PSU & 0.083° × 0.083° & daily & E.U. Copernicus Marine Service Information${^b}{^c}$\\
Mixed Layer Depth & mld & m & 0.083° × 0.083° & daily & E.U. Copernicus Marine Service Information${^b}{^c}$\\
8-day average Chlorophyll-a & chl8 & mg/m${^3}$ & 0.036° × 0.036° & 8-days composite & ERDDAP/NOAA${^d}$\\
\addlinespace
Log10 8-day average Chlorophyll-a & log10\_chl8 & $log_{10}$(chl8) & 0.036° × 0.036° & 8-days composite & Derived from ERDDAP/NOAA${^d}$\\
Eddy Kinetic Energy & eke & m/s & 0.2° × 0.2° & daily & IMOS/AODN${^e}$\\
Climatic mean EKE & clim\_eke\_mean & mean(m/s)/season/10 years & 0.2° × 0.2° & seasonal climatology & Derived from IMOS/AODN${^e}$\\
Climatic standard deviation EKE & clim\_eke\_sd & sd(m/s)/season/10 years & 0.2° × 0.2° & seasonal climatology & Derived from IMOS/AODN${^e}$\\
Climatic mean SSTgrad & clim\_sst\_grad & mean(Δ°Celsius)/season/10 years & 0.25° × 0.25° & seasonal climatology & Derived from Huang et al. (2021)${^a}$\\
\addlinespace
Bathymetry & bat & m & 0.004° × 0.004° & NA & GEBCO${^f}$\\
Slope & slope & degrees (°) & 0.004° × 0.004° & NA & Derived from GEBCO${^f}$\\
Distance from coast & dist\_coast & km & NA & NA & Calculated in R (see main text)\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{a} NOAA OISST: (https://psl.noaa.gov/data/gridded/data.noaa.oisst.v2.highres.html); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{b} 2011--2019: GLOBAL\_MULTIYEAR\_PHY\_001\_030 (https://doi.org/10.48670/moi-00021); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{c} 2020--2021: GLOBAL\_ANALYSIS\_FORECAST\_PHY\_001\_024 (https://doi.org/10.48670/moi-00016); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{d} ERDDAP/NOAA (https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1chla8day.html); }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{e} IMOS/AODN (https://portal.aodn.org.au/search), 'OceanCurrent - Gridded sea level anomaly - Delayed mode' data set; }\\
\multicolumn{6}{l}{\rule{0pt}{1em}\textsuperscript{f} GEBCO (https://www.gebco.net/data\_and\_products/gridded\_bathymetry\_data/)}\\
\end{tabular}
\end{table}

\end{landscape}

\newpage

\begin{table}

\caption{\label{tab:table2} }
\centering
\begin{tabular}[t]{>{}lll>{\raggedright\arraybackslash}p{11cm}l}
\toprule
 &  & Model & Model specification & BIC\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{\textbf{Bernoulli}}}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Summer}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad & 332.69\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sss & 319.34\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Autumn}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad & 1184.29\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sst & 1147.94\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Winter}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad + clim\_sst\_grad & 1089.43\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst & 1062.98\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Spring}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + dist\_coast + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad + clim\_eke\_mean & 2345.6\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sst & 2306.7\\
\cmidrule{1-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{\textbf{Negative Binomial}}}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Summer}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad & 1518.96\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + sss & 1505.13\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Autumn}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad & -Inf (n.f.)\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & bat + chl + sst + mld + clim\_sst\_grad & 3082.14\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Winter}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + eke + sst + sss + mld + sst\_grad + clim\_sst\_grad & 3262.45\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst & 3234.21\\
\cmidrule{2-5}
\addlinespace[0.3em]
\multicolumn{5}{l}{\textit{Spring}}\\
\hspace{1em}\hspace{1em} &  & a & bat + slope + dist\_coast + eke + chl + sst + sss + mld + sst\_grad + clim\_sst\_grad + clim\_eke\_mean & 5854.32\\
\cmidrule{3-5}
\hspace{1em}\hspace{1em} &  & b & sst + clim\_eke\_mean & 5824.38\\
\bottomrule
\end{tabular}
\end{table}

\newpage

\begin{landscape}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../EDA/Fig1-study-area-raw-counts}

}

\caption{ }\label{fig:fig1-study-area}
\end{figure}

\end{landscape}

\newpage

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../results/Fig4_partial-plots-spring}

}

\caption{ }\label{fig:fig2-partial-plots}
\end{figure}

\newpage

\begin{figure}

{\centering \includegraphics[width=1.1\linewidth]{../results/Fig2_point-predictions}

}

\caption{ }\label{fig:fig3-point-pred}
\end{figure}

\newpage

\begin{landscape}

\begin{figure}

{\centering \includegraphics[width=1.1\linewidth]{../results/Fig3_probability-maps-spring}

}

\caption{ }\label{fig:fig4-prob-maps}
\end{figure}

\newpage

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{../results/Fig5_spp-profiles-spring}

}

\caption{ }\label{fig:fig5-spp-profiles}
\end{figure}

\end{landscape}

\bibliographystyle{unsrt}
\bibliography{../references.bib}

\end{document}