### 2024-03-06 ms_PROOCE/ms_PROOCE_supplementary-material.Rmd

---
title: "Supplementary Material"
subtitle: "**Seabird assemblages are linked to the major western boundary current off eastern Australia**"
author: "Daudt, N.W.; Woehler, E.J.; Schofield, M.R.; Smith, R.O.; Bugoni, L.; Rayment, W.J."
date: "`r Sys.Date()`"
output:
  pdf_document:
    keep_tex: true
bibliography: ../references.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/elsevier-harvard.csl
classoption: a4paper
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include = FALSE, echo = FALSE}
library(readr)
library(dplyr)
library(purrr)
library(stringr)
library(gt)
```

This document is a supplementary file for @daudt2024. Briefly, here you will find:

-   **List of packages** (and version) used for data wrangling, visualisation and analyses, and their full references;
-   **Table 1**: Summary of sampling effort by voyage;
-   **Fig. 1**: Number of occurrences, frequency of occurrence (FO), and numeric frequency (NF) for each recorded species by season;
-   **Figs. 2, 3**: Plots for choosing best RCP group number (`multifit`);
-   **Figs. 4, 5**: Residual plots from the best fitted models;
-   **Figs. 6, 7**: Partial plots for covariates;
-   **Figs. 8, 9**: Probability maps from seasonal predictions;
-   **Figs. 10, 11**: Species profiles for each season;
-   **Fig. 12**: Species-richness and sample-coverage curves.

\newpage

# List of packages

We used the following packages 'plyr' 1.8.8 [@plyr], 'dplyr' 1.1.2 [@dplyr], 'tidyr' 1.3.0 [@tidyr], 'readr' 2.1.4 [@readr], 'tibble' 3.2.1 [@tibble], 'lubridate' 1.9.2 [@lubridate], 'stringr' 1.5.0 [@stringr], 'purrr' 1.0.1 [@purrr], 'ggplot2' 3.4.2 [@ggplot2], 'ggspatial' 1.1.7 [@ggspatial], 'patchwork' 1.1.2 [@patchwork], 'RColorBrewer' 1.1-3 [@rcolorbrewer], 'rnaturalearth' 0.3.2 [@rnaturalearth], 'sp' 1.6-0 [@sp1; @sp2], 'sf' 1.0-8 [@sf], 'mapview' 2.11.0 [@mapview], 'raster' 3.5-21 [@raster], 'terra' 1.6-7 [@terra], 'rerddap' 1.0.2 [@rerddap], 'rerddapXtracto' 1.1.4 [@rerddapXtracto], 'hadsstR' [@hadsstr], 'corrplot' 0.92 [@corrplot2021], and the ones referenced in the main text.

The code is archived in an Open Science Framework repository in [@daudt_data_2023], where you can find a detailed walk-through.

# References {.unnumbered}

::: {#refs}
:::

\newpage

\blandscape

```{r table-S1, echo = FALSE}
read.csv("../results/tableS1-effort-summary.csv")  %>% 
  dplyr::select(-X) %>% 
  gt::gt() %>% 
  # Tidy-up column labels
  gt::cols_label(
    voyage = gt::md("**Voyage**"),
    date_start = gt::md("**Date start**"),
    date_end = gt::md("**Date end**"),
    lat_range = gt::md("**Latitudinal range**"), 
    lon_range = gt::md("**Longitudinal range**"),
    no_records = gt::md("**No. of records**"),
    no_birds = gt::md("**No. of birds**"),
    no_spp = gt::md("**No. of species**")) %>% 
  # Table title
  gt::tab_header(title = gt::md("Table 1. Summary of seabird sampling effort by voyage, off eastern Australia, during Australasian Seabird Group's ship-based surveys between 2016--2021. Start and finish dates and geographic ranges of each voyage, including the number of seabird records and the number of individuals and species recorded"))
```

\newpage

![Number of occurrences (nOCC), frequency of occurrence (FO) and numeric frequency (NF) of seabirds recorded off eastern Australia during Australasian Seabird Group's ship-based surveys between 2016--2021. The dashed line represents the number of occurrence thresholds (n = 6) each taxon had to match for its inclusion in the seasonal models (see *Methods* in the main text). Species are ordered from the largest to the lowest total number of occurrences.](../results/FigS1_spp-nOCC-FO-NF-seasons.png){width="180%"}

\elandscape

\newpage

![Multifit plot for Region of Common Profiles (RCP) for each seasonal presence-absence model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_1_multifit-Bernoulli.png){width="100%"}

\newpage

![Multifit plot for Region of Common Profiles (RCP) for each seasonal abundance (count) model, applied to seabirds off eastern Australia. The number of groups with the lowest BIC value indicates the best number of groups (assemblages) that describes the data. For each number of groups, we ran 100 models with random starting values to avoid getting stuck in an incorrect 'optima' (see *Methods* in the main text). The resulting plot also shows how many groups were 'empty' (colour scale) with 'five or less' or 'only one' sites assigned to an RCP, i.e. the model was fit with, say, 5 groups, but 3 of them had 'five or less' or 'only one' sites (grids) allocated to an RCP.](../results/FigS3_2_multifit-NegBin.png){width="100%"}

\newpage

![Best model residuals for each seasonal presence-absence Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_1_best-model-residuals-Bernoulli.png){width="100%"}

\newpage

![Best model residuals for each seasonal abundance (count) Region of Common Profile model, applied to seabirds off eastern Australia.](../results/FigS4_2_best-model-residuals-NegBin.png){width="100%"}

\newpage

![Partial plots for the retained covariates in the best seasonal models based on presence-absence data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_1_partial-plots-Bernoulli.png){width="90%"}

\newpage

\blandscape

![Partial plots for the retained covariates in the best seasonal models based on abundance data. The plot shows the probability of belonging to a Region of Common Profiles (RCP) against the environmental value. Refer to Table 1 in the main text for the environmental data acronyms.](../results/FigS6_2_partial-plots-NegBin.png){width="150%"}

\newpage

![Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from presence-absence models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_1_prob-pred-Bernoulli.png){width="165%"}

\newpage

![Predicted probability membership of for each seabird assemblage (Region of Common Profiles; RCP) and grid, off eastern Australia, from abundance (count) models. The central column, 'mean', corresponds to the point prediction and Bayesian bootstraped, lower and upper confidence intervals (CI), on its sides.](../results/FigS5_2_prob-pred-NegBin.png){width="165%"}

\newpage

![Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal presence-absence model. Values are the average and confidence intervals of probability of occurrence for each species, based on 1000 Bayesian bootstraps. Note the species on the Y-axis are ordered alphabetically, from bottom up.](../results/FigS7_1_spp-profiles-Bernoulli.png){width="145%"}

\newpage

![Species profiles for each assemblage (Region of Common Profiles; RCP) for each seasonal abundance (count) model. Values are the average and confidence intervals of predicted mean abundance for each species, based on 1000 Bayesian bootstraps. Values were log10-transformed to accommodate the high variation between species. Note the species on the Y-axis are ordered alphabetically, from bottom up.](../results/FigS7_2_spp-profiles-NegBin.png){width="145%"}

\elandscape

\newpage

![Diversity curve (alpha diversity) and sample coverage for each assemblage (Region of Common Profile; RCP) from each presence-absence seasonal model.](../results/FigS8_iNEXT.png){width="90%"}

\newpage